---
title: "Lasso (GDP)"
author: "Shuofan Zhang"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r, include=T, echo=F, warning=F, 	message = FALSE}
library("glmnet")
library("reshape2")
library("cellranger")
library("tidyverse")
library("knitr")
library("ggrepel")
library("pander")
library("kableExtra")
library("tseries")
library("urca")
source("/Users/stanza/documents/github/lasso/fun.R")
```

```{r, include=T, echo=F, warning=F, 	message = FALSE}
# read data
bpraw <- read.csv("/Users/stanza/documents/github/lasso/bpraw.csv") 
bptrans <- read.csv("/Users/stanza/documents/github/lasso/bptrans.csv", na.strings = NaN)
tcode <- read.csv("/Users/stanza/documents/github/lasso/tcode.csv", stringsAsFactors = F) 
short <- read.csv("/Users/stanza/documents/github/lasso/short.csv", stringsAsFactors = F) %>% unname()
long <- read.csv("/Users/stanza/documents/github/lasso/long.csv", stringsAsFactors = F) %>% unname()
idx <- apply(bpraw, 2, anyNA) %>% which() %>% unname()
colnames(bpraw) = short
colnames(bptrans) = short
bpraw <- bpraw[ , -idx]
bptrans <- bptrans[, -idx] %>% na.omit()
tcode <- tcode[-idx,] 
short <- short[, -idx] %>% t() %>% as.vector()
long <- long[, -idx] %>% t() %>% as.vector()
```

## Detailed description

Given the data set we use (Stock & Watson 2016), when thresh=1E-16, the difference in sum of squared errors between OLS and Lasso ($\lambda=0$) is 1.81e-06, while when thresh=1E-7, the difference is 0.005. But the difference when $\lambda$ is nonzero is not very significant, so to reduce the computational burden, we choose to use 1E-10 with the maxit (maximum number of passes over the data for all lambda values) to be 10^9. 

All the monthly data was aggregated into quarterly data. 

Each series was standardized (centered, sd=1) before put into regression.

118 series were transformed by log().

GDP growth rate is used as the dependent variable.
$$y_t = log(GDP_{t})$$
$$\Delta y_t = log(GDP_{t}) - log(GDP_{t-1})$$
```{r, include=T, message=F, warning=F, echo=F}
ggplot(bpraw, aes(x=(1:224), y=log(GDP))) + geom_line() +
  labs(caption = "Time plot of log(GDP).") +
  xlab("time")
diff(log(bpraw[,1])) %>% as.data.frame() %>% ggplot(aes(x=(1:223), y=.)) + geom_line() +
  labs(caption = "Time plot of differenced log(GDP).") +
  xlab("time") +
  ylab("First-differenced GDP")
```

### Lasso 1

```{r, include=T, message=F, warning=F, echo=F}
# clean data
maxlag.1 = 1
d.log.gdp <- bpraw[,1] %>% log() %>% diff() #%>% as.data.frame()# (2-224=223)
y.1 <- lag0(maxlag.1 , d.log.gdp) %>% scale() # dependent variable (3-224=222), done
x <- bpraw[, -1] %>% data.matrix()
x.tcode = tcode[-1] 
lambda.seq1 <- seq(0.5, 0, -0.0025)
trans.1 = "1 less differenced"
vecm.1 = FALSE
# log() transformation
x[ , x.tcode==5 | x.tcode==6] = log( x[ , x.tcode==5 | x.tcode==6] )
# first difference
xtem.1 = x[-1,] # tem (2-224=223)
xtem.1[ , x.tcode==6] <- diff( x[ , x.tcode==6] )
colnames(xtem.1)[x.tcode==2 | x.tcode==5 | x.tcode==6] = 
  paste0(colnames(xtem.1)[x.tcode==2 | x.tcode==5 | x.tcode==6], ".level")
x.1 = lag1(maxlag.1 , xtem.1) %>% scale() # explanatory variables (2-223=222), done
p.1 = dim(x.1)[2]
n.1 = dim(x.1)[1]
lambda.fix.1 = sqrt(log( p.1 ) / n.1 )
# fit model, collect data
lasso.1 = glmnet(x.1, y.1, alpha=1, thresh=1E-5, lambda= lambda.seq1, maxit = 10^9)
trace1 = trace.plot.table(lasso.1, lambda.seq1, lambda.fix.1)
fitted.1 = predict(lasso.1 , s=lambda.seq1[which.min(abs(lambda.seq1-lambda.fix.1))] , newx = x.1)
mse.1 = mean((y.1-fitted.1)^2)
mse.out1 = out.mse(x.1, y.1, lambda.seq1, p.1)
```


The regression being estimated is:
$$\Delta y_t = I(0)_{t-1}
+ I(1)_{t-1}
+ \Delta I(2)_{t-1}$$

1, 144 series were used as explanatory variables; 

2, I(2) series were first differenced, no change to I(0) and I(1) series;

3, all explanatory variables are lagged by 1 quarter;

4, after first difference and one lag, we have 222 observations (lost 2).

### Lasso 2

The regression being estimated is:
$$ 
\begin{split} 
\Delta y_t &= y_{t-1} \\
&+ \Delta y_{t-1}
+ \Delta y_{t-2} 
+ \Delta y_{t-3}
+ \Delta y_{t-4} \\
&+ I(0)_{t-1}
+ I(1)_{t-1}
+ \Delta I(2)_{t-1}
\end{split} 
$$ 

1, 149 series were used as explanatory variables, $log(GDP_{t-1}) \ \Delta log(GDP_{t-1}) \ \Delta log(GDP_{t-2}) \ \Delta log(GDP_{t-3}) \ \Delta log(GDP_{t-4})$ were added on the top of the Lasso 1;

2, I(2) series were first differenced;

3, all explanatory variables are lagged by 1 quarter;

4, after first difference and four lags, we have 219 observations (lost 5);


```{r, include=T, message=F, warning=F, echo=F}
# create data
maxlag.2=4
y.2 <- lag0(maxlag.2 , d.log.gdp) %>% scale() # dependent variable, done
log.gdp.lag1 <- lag1(maxlag.2 , log(bpraw[-1 , 1]))  
d.log.gdp.lags <- lags(maxlag.2 , d.log.gdp)   
xtem.2 <- lag1(maxlag.2 , xtem.1) # x.2 (5 to 223)
x.2 = cbind(log.gdp.lag1, d.log.gdp.lags, xtem.2) %>% scale() # explanatory variable, done
lambda.seq2 = lambda.seq1
p.2 = dim(x.2)[2]
n.2 = dim(x.2)[1]
lambda.fix.2 = sqrt(log( p.2 ) / n.2 )
trans.2 = "1 less differenced"
vecm.2 = TRUE
# fit model, collect data
lasso.2 = glmnet(x.2, y.2, alpha=1, thresh=1E-5, lambda= lambda.seq2, maxit = 10^9)
trace2 = trace.plot.table(lasso.2, lambda.seq2, lambda.fix.2)
fitted.2 = predict(lasso.2 , s=lambda.seq2[which.min(abs(lambda.seq2-lambda.fix.2))] , newx = x.2)
mse.2 = mean((y.2-fitted.2)^2)
mse.out2 = out.mse(x.2, y.2, lambda.seq2, p.2)
```

### Lasso 3

Some notes: in this section, the number of parameters exceeds the number of observations, but glmnet still works when $\lambda=0$ (why), and "lm" also works unless we set "singular.ok = FALSE".

The regression being estimated is:
$$ 
\begin{split} 
\Delta y_t &= y_{t-1} \\
&+ \Delta y_{t-1} + \Delta y_{t-2} + \Delta y_{t-3} + \Delta y_{t-4} \\
&+  I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} +  I(0)_{t-4} \\
&+  \Delta I(1)_{t-1} +  \Delta I(1)_{t-2} 
+  \Delta I(1)_{t-3} +  \Delta I(1)_{t-4} \\
&+  \Delta^2 I(2)_{t-1} + \Delta^2 I(2)_{t-2} 
+ \Delta^2 I(2)_{t-3} +  \Delta^2 I(2)_{t-4}
\end{split} 
$$ 



1, 581 (1+4*145) series were used as explanatory variables; 

I(1) series were first-differenced;

I(2) series were second-differenced;

2, all explanatory variables are now I(0) and lagged by 4 quarters;

3, after first difference, second difference and four lags, we have 218 observations (lost 6).

```{r, include=T, message=F, warning=F, echo=F}
# create data
maxlag.3 = 4
xtem.3 = lags(maxlag.3 , bptrans)
y.lag1.3 = lag1(maxlag.3 , log(bpraw[-(1:2) , 1]) )
x.3 = cbind(y.lag1.3 , xtem.3) %>% scale()
y.3 = lag0(maxlag.3 , bptrans[,1] ) %>% scale()
lambda.seq3 = seq(0.6, 0.001, -0.002995)
p.3 = dim(x.3)[2]
n.3 = dim(x.3)[1]
lambda.fix.3 = sqrt(log( p.3 ) / n.3 )
trans.3 = "All stationary"
vecm.3 = TRUE
# fit model, collect data
lasso.3 = glmnet(x.3, y.3, alpha=1, thresh=1E-5, lambda= lambda.seq3, maxit = 10^9)
trace3 = trace.plot.table(lasso.3, lambda.seq3, lambda.fix.3)
fitted.3 = predict(lasso.3 , s=lambda.seq3[which.min(abs(lambda.seq3-lambda.fix.3))] , newx = x.3)
mse.3 = mean((y.3-fitted.3)^2)
mse.out3 = out.mse(x.3, y.3, lambda.seq3, p.3)
```

### Lasso 4

The regression being estimated is:
$$ 
\begin{split} 
\Delta y_t &= y_{t-1} \\
&+ \Delta y_{t-1} + \Delta y_{t-2} + \Delta y_{t-3} \\
&+  I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} \\
&+  \Delta I(1)_{t-1} +  \Delta I(1)_{t-2} 
+  \Delta I(1)_{t-3} \\
&+  \Delta^2 I(2)_{t-1} + \Delta^2 I(2)_{t-2} 
+ \Delta^2 I(2)_{t-3} 
\end{split} 
$$ 



1, 436 (1+3*145) series were used as explanatory variables; 

I(1) series were first-differenced;

I(2) series were second-differenced;

2, all explanatory variables are now I(0) and lagged by 3 quarters;

3, after first difference, second difference and 3 lags, we have 219 observations (lost 5).

```{r, include=T, message=F, warning=F, echo=F}
# create data
maxlag.4 = 3
xtem.4 = lags(maxlag.4 , bptrans)
y.lag1.4 = lag1(maxlag.4 , log(bpraw[-(1:2) , 1]) )
x.4 = cbind(y.lag1.4 , xtem.4) %>% scale()
y.4 = lag0(maxlag.4 , bptrans[,1] ) %>% scale()
lambda.seq4 = seq(0.6, 0.001, -0.002995)
p.4 = dim(x.4)[2]
n.4 = dim(x.4)[1]
lambda.fix.4 = sqrt(log( p.4 ) / n.4 )
trans.4 = "All stationary"
vecm.4 = TRUE
# fit model, collect data
lasso.4 = glmnet(x.4, y.4, alpha=1, thresh=1E-5, lambda= lambda.seq4, maxit = 10^9)
trace4 = trace.plot.table(lasso.4, lambda.seq4, lambda.fix.4)
fitted.4 = predict(lasso.4 , s=lambda.seq4[which.min(abs(lambda.seq4-lambda.fix.4))] , newx = x.4)
mse.4 = mean((y.4-fitted.4)^2)
mse.out4 = out.mse(x.4, y.4, lambda.seq4, p.4)
```

### Lasso 5

The regression being estimated is:
$$ 
\begin{split} 
\Delta y_t &= y_{t-1} \\
&+ \Delta y_{t-1} + \Delta y_{t-2} \\
&+  I(0)_{t-1} +  I(0)_{t-2} \\
&+  \Delta I(1)_{t-1} +  \Delta I(1)_{t-2} \\
&+  \Delta^2 I(2)_{t-1} + \Delta^2 I(2)_{t-2} 
\end{split} 
$$ 

1, 291 (1+2*145) series were used as explanatory variables; 

I(1) series were first-differenced;

I(2) series were second-differenced;

2, all explanatory variables are now I(0) and lagged by 2 quarters;

3, after first difference, second difference and 2 lags, we have 219 observations (lost 5).

```{r, include=T, message=F, warning=F, echo=F}
# create data
maxlag.5 = 2
xtem.5 = lags(maxlag.5 , bptrans)
y.lag1.5 = lag1(maxlag.5 , log(bpraw[-(1:2) , 1]) )
x.5 = cbind(y.lag1.5 , xtem.5) %>% scale()
y.5 = lag0(maxlag.5 , bptrans[,1] ) %>% scale()
lambda.seq5 = seq(0.6, 0.001, -0.002995)
p.5 = dim(x.5)[2]
n.5 = dim(x.5)[1]
lambda.fix.5 = sqrt(log( p.5 ) / n.5 )
trans.5 = "All stationary"
vecm.5 = TRUE
# fit model, collect data
lasso.5 = glmnet(x.5, y.5, alpha=1, thresh=1E-5, lambda= lambda.seq5, maxit = 10^9)
trace5 = trace.plot.table(lasso.5, lambda.seq5, lambda.fix.5)
fitted.5 = predict(lasso.5 , s=lambda.seq5[which.min(abs(lambda.seq5-lambda.fix.5))] , newx = x.5)
mse.5 = mean((y.5-fitted.5)^2)
mse.out5 = out.mse(x.5, y.5, lambda.seq5, p.5)
```

### Lasso 6

The regression being estimated is:
$$ 
\begin{split} 
\Delta y_t &= y_{t-1} + \Delta y_{t-1} +  I(0)_{t-1} +  \Delta I(1)_{t-1} +  \Delta^2 I(2)_{t-1} 
\end{split} 
$$ 

1, 146 (1+145) series were used as explanatory variables; 

I(1) series were first-differenced;

I(2) series were second-differenced;

2, all explanatory variables are now I(0) and lagged by 1 quarters;

3, after first difference, second difference and 1 lag, we have 221 observations (lost 3).

```{r, include=T, message=F, warning=F, echo=F}
# create data
maxlag.6 = 1
xtem.6 = lags(maxlag.6 , bptrans)
y.lag1.6 = lag1(maxlag.6 , log(bpraw[-(1:2) , 1]) )
x.6 = cbind(y.lag1.6 , xtem.6) %>% scale()
y.6 = lag0(maxlag.6 , bptrans[,1] ) %>% scale()
lambda.seq6 = seq(0.6, 0.001, -0.002995)
p.6 = dim(x.6)[2]
n.6 = dim(x.6)[1]
lambda.fix.6 = sqrt(log( p.6 ) / n.6 )
trans.6 = "All stationary"
vecm.6 = TRUE
# fit model, collect data
lasso.6 = glmnet(x.6, y.6, alpha=1, thresh=1E-5, lambda= lambda.seq6, maxit = 10^9)
trace6 = trace.plot.table(lasso.6, lambda.seq6, lambda.fix.6)
fitted.6 = predict(lasso.6 , s=lambda.seq6[which.min(abs(lambda.seq6-lambda.fix.6))] , newx = x.6)
mse.6 = mean((y.6-fitted.6)^2)
mse.out6 = out.mse(x.6, y.6, lambda.seq6, p.6)
```

### Lasso 7

```{r, echo=F, message=F, warning=F}
maxlag.7=4
xtem.7 = xtem.1[ , x.tcode==2 | x.tcode==5 | x.tcode==6] 
xtem.7 = lag1(maxlag = maxlag.7, xtem.7[-1,])
colnames(xtem.7) = paste0(colnames(xtem.7), ".level")
x.7 = cbind(xtem.3, xtem.7) %>% scale()
y.7 = y.3
lambda.seq7 = seq(0.6, 0.001, -0.002995)
p.7 = dim(x.7)[2]
n.7 = dim(x.7)[1]
lambda.fix.7 = sqrt(log( p.7 ) / n.7 )
trans.7 = "Combined"
vecm.7 = TRUE
# fit model, collect data
lasso.7 = glmnet(x.7, y.7, alpha=1, thresh=1E-5, lambda= lambda.seq7, maxit = 10^9)
trace7 = trace.plot.table(lasso.7, lambda.seq7, lambda.fix.7)
fitted.7 = predict(lasso.7 , s=lambda.seq7[which.min(abs(lambda.seq7-lambda.fix.7))] , newx = x.7)
mse.7 = mean((y.7-fitted.7)^2)
mse.out7 = out.mse(x.7, y.7, lambda.seq7, p.7)
```

The regression being estimated is:
$$ 
\begin{split} 
\Delta y_t &=\Delta y_{t-1} + \Delta y_{t-2} + \Delta y_{t-3} + \Delta y_{t-4} \\
&+  I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} +  I(0)_{t-4} \\
&+  \Delta I(1)_{t-1} +  \Delta I(1)_{t-2} 
+  \Delta I(1)_{t-3} +  \Delta I(1)_{t-4} \\
&+  \Delta^2 I(2)_{t-1} + \Delta^2 I(2)_{t-2} 
+ \Delta^2 I(2)_{t-3} +  \Delta^2 I(2)_{t-4} \\
&+ I(1)_{t-1} 
+  \Delta I(2)_{t-1} 
\end{split} 
$$
1, 697 series were used as explanatory variables; 

I(1) series were first-differenced;

I(2) series were second-differenced;

2, all explanatory variables are now I(0) and lagged by 4 quarters;

3, after first difference, second difference and 4 lags, we have 218 observations (lost 6).

<!-- Lasso 8 gives the same results with lasso 7, omitted 

The regression being estimated is:
$$ 
\begin{split} 
\Delta y_t &=y_{t-1} \\
&+ \Delta y_{t-1} + \Delta y_{t-2} + \Delta y_{t-3} + \Delta y_{t-4} \\
&+  I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} +  I(0)_{t-4} \\
&+  \Delta I(1)_{t-1} +  \Delta I(1)_{t-2} 
+  \Delta I(1)_{t-3} +  \Delta I(1)_{t-4} \\
&+  \Delta^2 I(2)_{t-1} + \Delta^2 I(2)_{t-2} 
+ \Delta^2 I(2)_{t-3} +  \Delta^2 I(2)_{t-4} \\
&+ I(1)_{t-1} 
+  \Delta I(2)_{t-1}
\end{split} 
$$

The level of lag 1 of y ($y_{t-1}$) was added on the top of lasso 7.

1, 698 series were used as explanatory variables; 

I(1) series were first-differenced;

I(2) series were second-differenced;

2, all explanatory variables are now I(0) and lagged by 4 quarters;

3, after first difference, second difference and 4 lags, we have 218 observations (lost 6). -->

### Lasso 8 (no spread)

"tb6m_tb3m, GS1_tb3m, GS10_tb3m, CP_Tbill Spread" were omitted to prevent from multicollinearity.
"CP3FM" was added. (file path of data file is changed in "datain_all.m", 24th Oct)

This model is same with the lasso 1, only difference is the removed four I(0) series, and the added one I(1) series.

Among the four non-zero coefficients, two of them are I(1), the other two are I(0) according to ADF test. However, the linear combination (using the estimated coefficients) of the two I(1) is not I(0). 

```{r, echo=F, message=F, warning=F}
bpraw2 <- read.csv("/Users/stanza/documents/github/lasso/new/bpraw.csv") 
bptrans2 <- read.csv("/Users/stanza/documents/github/lasso/new/bptrans.csv", na.strings = NaN)
tcode2 <- read.csv("/Users/stanza/documents/github/lasso/new/tcode.csv", stringsAsFactors = F) 
short2 <- read.csv("/Users/stanza/documents/github/lasso/new/short.csv", stringsAsFactors = F) %>% unname()
long2 <- read.csv("/Users/stanza/documents/github/lasso/new/long.csv", stringsAsFactors = F) %>% unname()
# remove NA
idx2 <- apply(bpraw2, 2, anyNA) %>% which() %>% unname()
colnames(bpraw2) = short2
colnames(bptrans2) = short2
bpraw2 <- bpraw2[ , -idx2]
bptrans2 <- bptrans2[, -idx2] %>% na.omit()
tcode2 <- tcode2[-idx2,] 
short2 <- short2[, -idx2] %>% t() %>% as.vector()
long2 <- long2[, -idx2] %>% t() %>% as.vector()
inves.raw2 <- bpraw2
inves.tran2 <- bptrans2
inves.tcode2 <- tcode2
inves.short2 <- short2
inves.long2 <- long2
# remove "tb6m_tb3m, GS1_tb3m, GS10_tb3m, CP_Tbill Spread"
tcode2 <- tcode2[colnames(bpraw2) != "tb6m_tb3m" & colnames(bpraw2) != "GS1_tb3m" &
                    colnames(bpraw2) != "GS10_tb3m" & colnames(bpraw2) != "CP_Tbill Spread"]
short2 <- short2[colnames(bpraw2) != "tb6m_tb3m" & colnames(bpraw2) != "GS1_tb3m" &
                    colnames(bpraw2) != "GS10_tb3m" & colnames(bpraw2) != "CP_Tbill Spread"]
long2 <- long2[colnames(bpraw2) != "tb6m_tb3m" & colnames(bpraw2) != "GS1_tb3m" &
                    colnames(bpraw2) != "GS10_tb3m" & colnames(bpraw2) != "CP_Tbill Spread"]
bpraw2 <- bpraw2[ , colnames(bpraw2) != "tb6m_tb3m" & colnames(bpraw2) != "GS1_tb3m" &
                    colnames(bpraw2) != "GS10_tb3m" & colnames(bpraw2) != "CP_Tbill Spread"]
bptrans2 <- bptrans2[ , colnames(bptrans2) != "tb6m_tb3m" & colnames(bptrans2) != "GS1_tb3m" &
                    colnames(bptrans2) != "GS10_tb3m" & colnames(bptrans2) != "CP_Tbill Spread"]
# clean data
maxlag.8 = 1
y.8 =  bpraw2[,1] %>% log() %>% diff() %>% lag0(maxlag = maxlag.8, .) %>% scale() # (2-224=223)
x8 <- bpraw2[, -1] %>% data.matrix()
x.tcode2 = tcode2[-1] 
lambda.seq8 <- seq(0.5, 0, -0.0025)
trans.8 = "1 less differenced"
# log() transformation
x8[ , x.tcode2==5 | x.tcode2==6] = log( x8[ , x.tcode2==5 | x.tcode2==6] )
# first difference
xtem.8 = x8[-1,] # tem (2-224=223)
xtem.8[ , x.tcode2==6] <- diff( x8[ , x.tcode2==6] )
colnames(xtem.8)[x.tcode2==2 | x.tcode2==5 | x.tcode2==6] = 
  paste0(colnames(xtem.8)[x.tcode2==2 | x.tcode2==5 | x.tcode2==6], ".level")
x.8 = lag1(maxlag.8 , xtem.8) %>% scale() # explanatory variables (2-223=222), done
p.8 = dim(x.8)[2]
n.8 = dim(x.8)[1]
lambda.fix.8 = sqrt(log( p.8 ) / n.8 )
# fit model, collect data
lasso.8 = glmnet(x.8, y.8, alpha=1, thresh=1E-5, lambda= lambda.seq8, maxit = 10^9)
trace8 = trace.plot.table(lasso.8, lambda.seq8, lambda.fix.8)
fitted.8 = predict(lasso.8 , s=lambda.seq8[which.min(abs(lambda.seq8-lambda.fix.8))] , newx = x.8)
mse.8 = mean((y.8-fitted.8)^2)
mse.out8 = out.mse(x.8, y.8, lambda.seq8, p.8)
```

### Lasso 9

```{r, echo=F, message=F, warning=F}
maxlag.9 = 4
xtem.9 = bptrans2 %>% lags(maxlag = maxlag.9, .)
x.9 = lag1(xtem.8[-1, x.tcode2==2 | x.tcode2==5 | x.tcode2==6], maxlag = maxlag.9) %>% cbind(., xtem.9) %>% scale()
y.9 = lag0(y.8, maxlag = maxlag.9) %>% scale()
lambda.seq9 <- seq(0.5, 0, -0.0025)
p.9 = dim(x.9)[2]
n.9 = dim(x.9)[1]
lambda.fix.9 = sqrt(log( p.9 ) / n.9 )
# fit model, collect data
lasso.9 = glmnet(x.9, y.9, alpha=1, thresh=1E-5, lambda= lambda.seq9, maxit = 10^9)
trace9 = trace.plot.table(lasso.9, lambda.seq9, lambda.fix.9)
fitted.9 = predict(lasso.9 , s=lambda.seq9[which.min(abs(lambda.seq9-lambda.fix.9))] , newx = x.9)
mse.9 = mean((y.9-fitted.9)^2)
mse.out9 = out.mse(x.9, y.9, lambda.seq9, p.9)
```


### Investigation of the non-zero coefficients

```{r, echo=F, message=F, warning=F, eval=FALSE}
# investigate the non-zero series in all lasso models
inves.1 = x.1[, colnames(x.1) %in% (trace1$coef[,2] %>% as.character())]
inves.2 = x.2[, colnames(x.2) %in% (trace2$coef[,2] %>% as.character())]
inves.3 = x.3[, colnames(x.3) %in% (trace3$coef[,2] %>% as.character())]
inves.4 = x.4[, colnames(x.4) %in% (trace4$coef[,2] %>% as.character())]
inves.5 = x.5[, colnames(x.5) %in% (trace5$coef[,2] %>% as.character())]
inves.6 = x.6[, colnames(x.6) %in% (trace6$coef[,2] %>% as.character())]
inves.7 = x.7[, colnames(x.7) %in% (trace7$coef[,2] %>% as.character())]
inves.8 = x.8[, colnames(x.8) %in% (trace8$coef[,2] %>% as.character())]
inves.9 = x.9[, colnames(x.9) %in% (trace9$coef[,2] %>% as.character())]
AA = rbind(adf.auto(inves.1, criteria = "AIC") , adf.auto(inves.2, criteria = "AIC") , 
      adf.auto(inves.5, criteria = "AIC") , adf.auto(inves.6, criteria = "AIC") , adf.auto(inves.7, criteria = "AIC") , adf.auto(inves.8, criteria = "AIC") ,
      adf.auto(inves.9, criteria = "AIC")) %>% unique() 
BB = rbind(adf.auto(inves.1, criteria = "BIC") , adf.auto(inves.2, criteria = "BIC") , 
      adf.auto(inves.5, criteria = "BIC") , adf.auto(inves.6, criteria = "BIC") , adf.auto(inves.7, criteria = "BIC") , adf.auto(inves.8, criteria = "BIC") ,
      adf.auto(inves.9, criteria = "BIC")) %>% unique() 
AABB = left_join(AA, BB, by = "Series")
AABB = AABB[,-3]
AABB = AABB[,c(1,2,3,4,6,5)]
colnames(AABB) = c("Series", "AIC", "AIC lags", "BIC", "BIC lags", "Type")
```

### Investigation of the all variables

```{r, echo=F, message=F, warning=F}
sw2 = inves.raw2[,inves.tcode2==6]
adf.auto(sw2, criteria = "AIC")
inves.raw2[,inves.tcode2==5 | inves.tcode2==6] = 
  inves.raw2[,inves.tcode2==5 | inves.tcode2==6] %>% log()
temp = apply(inves.raw2[,inves.tcode2==6], 2, diff)
temp = rbind(NA, temp)
inves.raw2[,inves.tcode2==6] = temp
adf.aic = adf.auto(inves.raw2, criteria = "AIC")
adf.bic = adf.auto(inves.raw2, criteria = "BIC")
sw = data.frame(Series = colnames(inves.raw2), SW = character(length = 146), 
                  stringsAsFactors = FALSE)
sw$SW[inves.tcode2==1] = "I(0)"
sw$SW[inves.tcode2==2] = "I(1)"
sw$SW[inves.tcode2==5] = "I(1)"
sw$SW[inves.tcode2==6] = "I(2)"
aabbss <- left_join(sw, adf.aic, by = "Series") %>% left_join(., adf.bic, by = "Series")
aabbss <- aabbss[,-4]
aabbss <- aabbss[,c(1,2,3,5,6,4,7)]
colnames(aabbss) = c("Series", "SW", "AIC", "BIC", "Type", "AIC lags", "BIC lags")
tt = aabbss[inves.tcode2==6,]
tt$AIC[tt$AIC=="I(1)"] <- "I(2)"
tt$BIC[tt$BIC=="I(1)"] <- "I(2)"
tt$AIC[tt$AIC=="I(0)"] <- "I(1)"
tt$BIC[tt$BIC=="I(0)"] <- "I(1)"
aabbss[inves.tcode2==6,] = tt
saveRDS(aabbss, file = "/Users/stanza/documents/github/lasso/compare.rds")
```

## Graphs

```{r echo=F, message=F, warning=F}
# plot
plot1 = trace1$plot + ylim(-1, 1) + labs(subtitle = "Lasso 1, y is truncated to (-1, 1)")
plot1

plot2 = trace2$plot + ylim(-1, 1) + labs(subtitle = "Lasso 2, y is truncated to (-1, 1)")
plot2

plot3 = trace3$plot
plot3 + labs(subtitle = "Lasso 3")

plot4 = trace4$plot
plot4 + labs(subtitle = "Lasso 4")

plot5 = trace5$plot
plot5 + labs(subtitle = "Lasso 5")

plot6 = trace6$plot
plot6 + labs(subtitle = "Lasso 6")

plot7 = trace7$plot
plot7 + labs(subtitle = "Lasso 7")

plot8 = trace8$plot
plot8 + ylim(-1, 1) + labs(subtitle = "Lasso 8, y is truncated to (-1, 1)")

plot9 = trace9$plot
plot9 + labs(subtitle = "Lasso 9")
```



```{r, echo=F, message=F, warning=F}
# summary table create
variable = c(trace1$coef[,2] %>% as.character(), 
             trace2$coef[,2] %>% as.character(), 
             trace3$coef[,2] %>% as.character(), 
             trace4$coef[,2] %>% as.character(), 
             trace5$coef[,2] %>% as.character(), 
             trace6$coef[,2] %>% as.character(),
             trace7$coef[,2] %>% as.character(),
             trace8$coef[,2] %>% as.character(),
             trace9$coef[,2] %>% as.character()) %>% unique() %>% 
  as.data.frame(stringsAsFactors = FALSE)
colnames(variable) = "variable"
sum.table = left_join(variable, trace1$coef[,2:3]) %>% 
  left_join(., trace2$coef[,2:3], by = "variable") %>%
  left_join(., trace3$coef[,2:3], by = "variable") %>%
  left_join(., trace4$coef[,2:3], by = "variable") %>%
  left_join(., trace5$coef[,2:3], by = "variable") %>%
  left_join(., trace6$coef[,2:3], by = "variable") %>% 
  left_join(., trace7$coef[,2:3], by = "variable") %>% 
  left_join(., trace8$coef[,2:3], by = "variable") %>% 
  left_join(., trace9$coef[,2:3], by = "variable") %>% 
  as.data.frame(stringsAsFactors = FALSE)
sum.table[ , -1] = round(sum.table[ , -1] , digits = 6)
colnames(sum.table) = c("" , "Lasso 1" , "Lasso 2" , "Lasso 3" , 
                        "Lasso 4" , "Lasso 5" , "Lasso 6", "Lasso 7", "Lasso 8", "Lasso 9")

info.table = c("0" , "0" , "1", "0" , "0" , "1" , "1") %>% 
  cbind(c("1" , "4" , "1" , "0" , "0" , "1" , "1")) %>% 
  cbind(c("1" , "4" , "4" , "4" , "4" , "0" , "0")) %>% 
  cbind(c("1" , "3" , "3" , "3" , "3" , "0" , "0")) %>% 
  cbind(c("1" , "2" , "2" , "2" , "2" , "0" , "0")) %>% 
  cbind(c("1" , "1" , "1" , "1" , "1" , "0" , "0")) %>% 
  cbind(c("0" , "4" , "4" , "4" , "4" , "1" , "1")) %>% 
  cbind(c("0" , "0" , "1", "0" , "0" , "1" , "1")) %>% 
  cbind(c("0" , "0" , "4", "4" , "4" , "1" , "1")) %>%
  rbind(c(n.1 , n.2 , n.3 , n.4 , n.5 , n.6 , n.7 , n.8, n.9)) %>%
  rbind(c(p.1 , p.2 , p.3 , p.4 , p.5 , p.6 , p.7 , p.8, p.9)) %>%
  rbind(round(c(lambda.fix.1 , lambda.fix.2 , lambda.fix.3 , 
          lambda.fix.4 , lambda.fix.5 , lambda.fix.6 , 
          lambda.fix.7 , lambda.fix.8, lambda.fix.9), digits = 4)) %>% 
  rbind(round(c(mse.1 , mse.2 , mse.3 , mse.4 , mse.5 , mse.6 , mse.7 , mse.8, mse.9) , digits=5)) %>% 
  rbind(round(c(mse.out1 , mse.out2 , mse.out3 , mse.out4 , mse.out5 , mse.out6 , 
                mse.out7 , mse.out8, mse.out9) , digits=5)) %>% 
  cbind(c("y.lag1", "D.y.lags", "I(0).lags" ,"D.I(1).lags",
                     "D2.I(2).lags", "I(1).lags", "D.I(2).lags", 
               "#obs" , "#regressors" , "Fixed Lambda" , "MSE.in", "MSE.out") , .) %>% 
  as.data.frame(stringsAsFactors = FALSE)
colnames(info.table) = c("" , "Lasso 1" , "Lasso 2" , "Lasso 3" , 
                        "Lasso 4" , "Lasso 5" , "Lasso 6", "Lasso 7", "Lasso 8", "Lasso 9") 

saveRDS(sum.table, file = "/Users/stanza/documents/github/lasso/sum_table.rds")
saveRDS(info.table, file = "/Users/stanza/documents/github/lasso/info_table.rds")
```


## Data transformation table

```{r, results='asis', echo=F, warning=F, message=F}
t.data <- tibble(tcode, short, long) 

t1 <- t.data[t.data$tcode==1, ] 
t2 <- t.data[t.data$tcode==2, ] 
t3 <- t.data[t.data$tcode==3, ] 
t4 <- t.data[t.data$tcode==4, ] 
t5 <- t.data[t.data$tcode==5, ] 
t6 <- t.data[t.data$tcode==6, ] 

t1[,1] <- c("I(0)") 
t2[,1] <- c("I(1)")
t5[,1] <- c("log, I(1)")
t6[,1] <- c("log, I(2)")

pander(t1, justify = 'left', caption = "Number of series with 'No-transformation' is 12",
       split.cells = c(5, 5, 40)) 
pander(t2, justify = 'left', caption = "Number of 'First-differenced' series is 15",
       split.cells = c(5, 5, 40)) 
pander(t5, justify = 'left', caption = "Number of 'First-differenced in logs' series is 86",
       split.cells = c(5, 5, 40)) 
pander(t6, justify = 'left', caption = "Number of 'Second-differenced in logs' series is 32",
       split.cells = c(5, 5, 40)) 

```



```{r, echo=F, message=F, warning=F, eval=F}
# lasso 8, deleted
maxlag.8=4
x.8 = cbind("GDP.lag1.level"=y.lag1.3, xtem.3, xtem.7) %>% scale()
y.8 = y.3
lambda.seq8 = seq(0.6, 0.001, -0.002995)
p.8 = dim(x.8)[2]
n.8 = dim(x.8)[1]
lambda.fix.8 = sqrt(log( p.8 ) / n.8 )
trans.8 = "Combined"
vecm.8 = TRUE
# fit model, collect data
lasso.8 = glmnet(x.8, y.8, alpha=1, thresh=1E-5, lambda= lambda.seq8, maxit = 10^9)
trace8 = trace.plot.table(lasso.8, lambda.seq8, lambda.fix.8)
fitted.8 = predict(lasso.8 , s=lambda.seq8[which.min(abs(lambda.seq8-lambda.fix.8))] , newx = x.8)
mse.8 = mean((y.8-fitted.8)^2)
mse.out8 = out.mse(x.8, y.8, lambda.seq8, p.8)
```




```{r, echo=F, message=F, warning=F, eval=F}
## C.V.
cv7 = cv.glmnet(x.7, y.7, lambda = lambda.seq7,  alpha=1)
coef7 = coef(cv7, s="lambda.min") %>%
     as.matrix() %>% as.data.frame(stringsAsFactors = F, row.names = NULL)
 colnames(coef7) = "coef"
no.coef7 = subset(coef7, coef != 0)
```


```{r, eval=F, echo=F}
# experiment with levels only
test.x <- bpraw %>% data.matrix()
test.x[ , tcode==5 | tcode==6] = log(test.x[ , tcode==5 | tcode==6])
test.xtem = test.x[-1,]
test.xtem[ , tcode==6] <- diff(test.x[ , tcode==6])
test.x1 = lag1(maxlag.1, test.xtem) %>% scale()
test.x1 = test.x1[ , tcode==5 | tcode==6]
lambda.test = sqrt(log( dim(test.x1)[2] ) / dim(test.x1)[1] )
lasso.test = glmnet(test.x1, y.1, alpha=1, thresh=1E-5, lambda= lambda.seq1, maxit = 10^9)
test.trace = trace.plot.table(lasso.test, lambda.seq1, lambda.test)
test.trace$plot
test.trace$table
```


```{r, echo=FALSE, eval=FALSE}
`Ch. Inv/GDP.lag1` = x.8[, colnames(x.8) == "Ch. Inv/GDP.lag1"]
`IP: Dur gds materials.level.lag1` = x.1[, colnames(x.1) == "IP: Dur gds materials.level.lag1"]
`NAPM:ORD.lag1` = x.8[, colnames(x.8) == "NAPM:ORD.lag1"]
`CP_Tbill Spread.lag1` = x.1[, colnames(x.1) == "CP_Tbill Spread.lag1"]
`Ex rate: major.level.lag1` = x.1[, colnames(x.1) == "Ex rate: major.level.lag1"]
`Cons. Expectations.lag1` = x.8[, colnames(x.8) == "Cons. Expectations.lag1"]
`UnfOrders(DurGds).level.lag1` = x.2[, colnames(x.2) == "UnfOrders(DurGds).level.lag1"]
`Cons:Svc.lag1` = x.4[, colnames(x.4) == "Cons:Svc.lag1"]
`FixedInv:Res.lag1` = x.6[, colnames(x.6) == "FixedInv:Res.lag1"]
`IP: Nondur gds materials.lag1` = x.6[, colnames(x.6) == "IP: Nondur gds materials.lag1"]
`Emp:SlackWk.lag1` = x.7[, colnames(x.7) == "Emp:SlackWk.lag1"]
`CPI_LFE.lag2` = x.5[, colnames(x.5) == "CPI_LFE.lag2"]
`S&P 500.lag1` = x.6[, colnames(x.6) == "S&P 500.lag1"]
`DJIA.lag1` = x.6[, colnames(x.6) == "DJIA.lag1"]
`Emp: Const.level.lag1` = x.8[, colnames(x.8) == "Emp: Const.level.lag1"]
# ADF test
`Ch. Inv/GDP.lag1` %>% adf.auto() %>% pander(caption="Ch. Inv/GDP.lag1")
`IP: Dur gds materials.level.lag1` %>% adf.auto() %>% pander(caption="IP: Dur gds materials.level.lag1")
`NAPM:ORD.lag1` %>% adf.auto() %>% pander(caption = "NAPM:ORD.lag1")
`CP_Tbill Spread.lag1` %>% adf.auto() %>% pander(caption = "CP_Tbill Spread.lag1")
`Ex rate: major.level.lag1` %>% adf.auto() %>% pander(caption = "Ex rate: major.level.lag1")
`Cons. Expectations.lag1` %>% adf.auto() %>% pander(caption = "Cons. Expectations.lag1")
`UnfOrders(DurGds).level.lag1` %>% adf.auto() %>% pander(caption= "UnfOrders(DurGds).level.lag1")
`Cons:Svc.lag1` %>% adf.auto() %>% pander(caption = "Cons:Svc.lag1")
`FixedInv:Res.lag1` %>% adf.auto() %>% pander(caption="FixedInv:Res.lag1")
`IP: Nondur gds materials.lag1` %>% adf.auto() %>% pander(caption="IP: Nondur gds materials.lag1")
`Emp:SlackWk.lag1` %>% adf.auto() %>% pander(caption="Emp:SlackWk.lag1")
`CPI_LFE.lag2` %>% adf.auto() %>% pander(caption="CPI_LFE.lag2")
`S&P 500.lag1` %>% adf.auto() %>% pander(caption="S&P 500.lag1")
`DJIA.lag1` %>% adf.auto() %>% pander(caption="DJIA.lag1")
`Emp: Const.level.lag1` %>% adf.auto() %>% pander(caption="Emp: Const.level.lag1")
```

```{r, echo=F, message=F, warning=F, eval=F}
ggplot(data=as.data.frame(x.8) , aes(x=(1:222), y=`Emp: Const.lag1`)) + geom_line() + xlab("time")


ggplot(data=as.data.frame(x.8) , aes(x=(1:222), y=`Cons. Expectations.lag1`)) + geom_line() + xlab("time")


ggplot(data=as.data.frame(x.8) , aes(x=(1:222), y=`NAPM:ORD.lag1`)) + geom_line() + xlab("time")


ggplot(data=as.data.frame(x.8) , aes(x=(1:222), y=`Ch. Inv/GDP.lag1`)) + geom_line() + xlab("time")
```




