---
title: "Lasso (Inflation)"
author: "Shuofan Zhang"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r, include=T, echo=F, warning=F, 	message = FALSE}
library("glmnet")
library("reshape2")
library("cellranger")
library("tidyverse")
library("knitr")
library("ggrepel")
library("pander")
library("kableExtra")
library("tseries")
source("/Users/stanza/documents/github/lasso/fun.R")
```

```{r, include=T, echo=F, warning=F, 	message = FALSE}
# read data
bpraw <- read.csv("/Users/stanza/documents/github/lasso/bpraw.csv") 
bptrans <- read.csv("/Users/stanza/documents/github/lasso/bptrans.csv", na.strings = NaN)
tcode <- read.csv("/Users/stanza/documents/github/lasso/tcode.csv", stringsAsFactors = F) 
short <- read.csv("/Users/stanza/documents/github/lasso/short.csv", stringsAsFactors = F) %>% unname()
long <- read.csv("/Users/stanza/documents/github/lasso/long.csv", stringsAsFactors = F) %>% unname()
# need to remove the copyright symbol from the long.csv
idx <- apply(bpraw, 2, anyNA) %>% which() %>% unname()
colnames(bpraw) = short
colnames(bptrans) = short
bpraw <- bpraw[ , -idx]
bptrans <- bptrans[, -idx] %>% na.omit()
tcode <- tcode[-idx,] 
short <- short[, -idx] %>% t() %>% as.vector()
long <- long[, -idx] %>% t() %>% as.vector()
```

## Detailed description

The inflation rate is used as the dependent variable.
$$y_t = log(CPI_{t}) - log(CPI_{t-1})$$
$$\Delta y_t = log(CPI_{t}) - 2*log(CPI_{t-1}) + log(CPI_{t-2})$$

```{r, include=T, message=F, warning=F, echo=F}
ggplot(bpraw, aes(x=(1:224), y=log(CPI))) + geom_line() +
  labs(caption = "Time plot of log(CPI).") +
  xlab("time")
```

```{r, include=T, message=F, warning=F, echo=F}
diff(log(bpraw[,101])) %>% as.data.frame() %>% ggplot(aes(x=(1:223), y=.)) + geom_line() +
  labs(caption = "Time plot of first-differenced log(CPI) (i.e. inflation rate).") +
  xlab("time") +
  ylab("First-differenced CPI (inflation rate)")


diff(log(bpraw[,101])) %>% diff() %>% as.data.frame() %>% ggplot(aes(x=(1:222), y=.)) + geom_line() + labs(caption = "Time plot of second-differenced log(CPI) (first-differenced inflation rate") + xlab("time")

 
inflation <- bpraw[,101] %>% log() %>% diff() 
adf.inflation <- adf.test(inflation)
pander(adf.inflation)
```

ADF test suggests inflation rate is I(1) at 10% significance level.


### Lasso

The regression being estimated is:
$$ 
\begin{split} 
\Delta y_t &=y_{t-1} \\
&+ \Delta y_{t-1} + \Delta y_{t-2} + \Delta y_{t-3} + \Delta y_{t-4} \\
&+  I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} +  I(0)_{t-4} \\
&+  \Delta I(1)_{t-1} +  \Delta I(1)_{t-2} 
+  \Delta I(1)_{t-3} +  \Delta I(1)_{t-4} \\
&+  \Delta^2 I(2)_{t-1} + \Delta^2 I(2)_{t-2} 
+ \Delta^2 I(2)_{t-3} +  \Delta^2 I(2)_{t-4} \\
&+ I(1)_{t-1} 
+  \Delta I(2)_{t-1}
\end{split} 
$$

713 series are used as explanatory variables. 

4 lags of I(0), including first-differenced I(1) and second-differenced I(2), 4*145=580; (lags of $\Delta y_{t-1}, \Delta y_{t-2} , \Delta y_{t-3} , \Delta y_{t-4}$ are included)

1 lag of I(1), including first-differenced I(2), 133. ($y_{t-1}$ is included)

```{r, echo=F, message=F, warning=F}
maxlag.1=4
d.inflation <- diff(inflation)
y.1 <- lag0(maxlag.1 , d.inflation) %>% scale() # 221 obs, y is done
x.trans = lags(maxlag.1 , bptrans)
I1.log = bpraw[, tcode==5] %>% log()
I1 = cbind(bpraw[, tcode==2], I1.log)
d.I2 = bpraw[, tcode==6] %>% log() %>% apply(2, diff) %>% as.data.frame()
xtem = lag1(maxlag.1, I1[-(1:2),]) %>% cbind(lag1(maxlag.1, d.I2[-1,]))
colnames(xtem) = paste0(colnames(xtem), ".level")
x.1 <- cbind(x.trans , xtem) %>% scale() # x is done
lambda.seq.1 = seq(0.6, 0.001, -0.002995)
p.1 = dim(x.1)[2]
n.1 = dim(x.1)[1]
lambda.fix.1 = sqrt(log( p.1 ) / n.1 )
trans.1 = "Combined"
# fit model, collect data
lasso.1 = glmnet(x.1, y.1, alpha=1, thresh=1E-10, lambda= lambda.seq.1, maxit = 10^9)
trace1 = trace.plot.table(lasso.1, lambda.seq.1, lambda.fix.1)
fitted.1 = predict(lasso.1 , s=lambda.seq.1[which.min(abs(lambda.seq.1-lambda.fix.1))] , newx = x.1)
mse.1 = mean((y.1-fitted.1)^2)
mse.out1 = out.mse(x.1, y.1, lambda.seq.1, p.1)
trace1$table
trace1$plot
```


```{r, echo=F, message=F, warning=F}
# investigate the level series
pced.gas.energy.level = d.I2[,17]
adf <- adf.test(pced.gas.energy.level)
pander(adf)
ggplot(data=as.data.frame(pced.gas.energy.level) , aes(x=(1:223), y=pced.gas.energy.level)) + geom_line() + ggtitle("First differenced PCED_GAS_ENERGY") + xlab("time")
```

ADF test suggests the "level" variable is actually not I(1).

Now, we use the "inflation rate" as the dependent variable. 

The regression being estimated is:
$$ 
\begin{split} 
Inflation_t &=log(CPI)_{t-1} \\
&+  I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} +  I(0)_{t-4} \\
&+  \Delta I(1)_{t-1} +  \Delta I(1)_{t-2} 
+  \Delta I(1)_{t-3} +  \Delta I(1)_{t-4} \\
&+  \Delta^2 I(2)_{t-1} + \Delta^2 I(2)_{t-2} 
+ \Delta^2 I(2)_{t-3} +  \Delta^2 I(2)_{t-4} \\
&+ I(1)_{t-1} 
+  \Delta I(2)_{t-1}
\end{split} 
$$

```{r, echo=F, message=F, warning=F}
maxlag.2=4
y.2 <- lag0(maxlag.2 , inflation[-1]) %>% scale() # 221 obs, y is done
log.cpi.lag1 = bpraw[-(1:2) , 101] %>% log() %>% lag1(maxlag.2, .)
x.trans.2 = lags(maxlag.2 , bptrans)
d.I2.2 = d.I2[, -26]
xtem.2 = lag1(maxlag.2, I1[-(1:2),]) %>% cbind(lag1(maxlag.2, d.I2.2[-1,]))
colnames(xtem.2) = paste0(colnames(xtem.2), ".level")
x.2 <- cbind(log.cpi.lag1, x.trans.2 , xtem.2) %>% scale() # x is done

lambda.seq.2 = seq(0.6, 0.001, -0.002995)
p.2 = dim(x.2)[2]
n.2 = dim(x.2)[1]
lambda.fix.2 = sqrt(log( p.2 ) / n.2 )
trans.2 = "Combined"
# fit model, collect data
lasso.2 = glmnet(x.2, y.2, alpha=1, thresh=1E-10, lambda= lambda.seq.2, maxit = 10^9)
trace2 = trace.plot.table(lasso.2, lambda.seq.2, lambda.fix.2)
fitted.2 = predict(lasso.2 , s=lambda.seq.2[which.min(abs(lambda.seq.2-lambda.fix.2))] , newx = x.2)
mse.2 = mean((y.2-fitted.2)^2)
mse.out2 = out.mse(x.2, y.2, lambda.seq.2, p.2)
trace2$table
trace2$plot
```




<!--  $y_{t-1}$ is excluded now. This part has been deleted. -->

```{r, echo=F, message=F, warning=F, eval=F}
x.2 <- cbind(x.trans , xtem[,-112]) %>% scale() # x is done
y.2 = y.1
p.2 = dim(x.2)[2]
n.2 = dim(x.2)[1]
lambda.seq.2 = lambda.seq.1
lambda.fix.2 = sqrt(log(p.2)/n.2)
lasso.2 = glmnet(x.2, y.1, alpha=1, thresh=1E-10, lambda= lambda.seq.1, maxit = 10^9)
trace2 = trace.plot.table(lasso.2, lambda.seq.2, lambda.fix.2)
fitted.2 = predict(lasso.2 , s=lambda.seq.2[which.min(abs(lambda.seq.2-lambda.fix.2))] , newx = x.2)
mse.2 = mean((y.2-fitted.2)^2)
mse.out2 = out.mse(x.2, y.2, lambda.seq.2, p.2)
trace2$table
trace2$plot
```
















