---
title: "SW and ADF test (GDP)"
author: "Shuofan Zhang"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r, include=T, echo=F, warning=F, 	message = FALSE}
library("glmnet")
library("reshape2")
library("cellranger")
library("tidyverse")
library("knitr")
library("ggrepel")
library("pander")
library("kableExtra")
library("tseries")
library("urca")
source("/Users/stanza/documents/github/lasso/fun.R")
```

## Data

Data: 145+1-5=141 series in total.

The "spread" series (difference between two I(1) series) are removed.

One of the I(1), "CP3FM" was omitted in the original data set, now is added.

log() is done.


```{r echo=F, message=F, warning=F}
bpraw <- read.csv("/Users/stanza/documents/github/lasso/new/bpraw.csv") 
bptrans <- read.csv("/Users/stanza/documents/github/lasso/new/bptrans.csv", na.strings = NaN)
tcode <- read.csv("/Users/stanza/documents/github/lasso/new/tcode.csv", stringsAsFactors = F) 
short <- read.csv("/Users/stanza/documents/github/lasso/new/short.csv", stringsAsFactors = F) %>% unname()
long <- read.csv("/Users/stanza/documents/github/lasso/new/long.csv", stringsAsFactors = F) %>% unname()
# remove NA
idx <- apply(bpraw, 2, anyNA) %>% which() %>% unname()
colnames(bpraw) = short
colnames(bptrans) = short
bpraw <- bpraw[ , -idx]
bptrans <- bptrans[, -idx] %>% na.omit()
tcode <- tcode[-idx,] 
short <- short[, -idx] %>% t() %>% as.vector()
long <- long[, -idx] %>% t() %>% as.vector()
# log()
bpraw[, tcode==5 | tcode==6] = log(bpraw[, tcode==5 | tcode==6])
# keep full set of data before remove "spread"
full.raw <- bpraw
full.tran <- bptrans
full.tcode <- tcode
full.short <- short
full.long <- long
# remove "BAA_GS10", "tb6m_tb3m, GS1_tb3m, GS10_tb3m, CP_Tbill Spread"
idx2 = which(colnames(bpraw) == "BAA_GS10" | 
               colnames(bpraw) == "tb6m_tb3m" | 
               colnames(bpraw) == "GS1_tb3m" | 
               colnames(bpraw) == "GS10_tb3m" | 
               colnames(bpraw) == "CP_Tbill Spread")
tcode <- tcode[-idx2]
short <- short[-idx2]
long <- long[-idx2]
bpraw <- bpraw[ , -idx2]
bptrans <- bptrans[ , -idx2]
```

## ADF test  

Step 1, ADF test to the 146 original series.

Step 2, mark "I(0)" variables as "I(0)".

Step 3, ADF test to the first-differenced 146 series.

Step 4, check for contradictions, found "PCED_RecServices" in AIC.

Step 5, mark "I(1)" variable as "I(2)" (including "PCED_RecServices").

Step 6, mark the rest as "I(1)".

Step 7, repeat the above 6 steps for both "AIC" and "BIC".


```{r, echo=F, message=F, warning=F}
t.aic = data.frame(integration = character(length = 146), stringsAsFactors = FALSE)
t.bic = data.frame(integration = character(length = 146), stringsAsFactors = FALSE)
t.sw = data.frame(integration = character(length = 146), stringsAsFactors = FALSE)
adf.aic = adf.auto(full.raw, criteria = "AIC")
adf.bic = adf.auto(full.raw, criteria = "BIC")
t.aic[adf.aic$Conclusion=="I(0)",] = "0"
t.bic[adf.bic$Conclusion=="I(0)",] = "0"
diff.full.raw = apply(full.raw , 2, diff)
adf.aic.diff = adf.auto(diff.full.raw, criteria = "AIC")
adf.aic = cbind(adf.aic, adf.aic.diff)
adf.bic.diff = adf.auto(diff.full.raw, criteria = "BIC")
adf.bic = cbind(adf.bic, adf.bic.diff)
#write.csv(adf.aic, "/Users/stanza/documents/github/lasso/adf_aic.csv")
#write.csv(adf.bic, "/Users/stanza/documents/github/lasso/adf_bic.csv")
t.aic[adf.aic[,6]=="I(1)",] = "2"
t.bic[adf.bic[,6]=="I(1)",] = "2"
t.aic[t.aic=="",] = "1"
t.bic[t.bic=="",] = "1"
t.sw[full.tcode==1,] = "0"
t.sw[full.tcode==2 | full.tcode==5,] = "1"
t.sw[full.tcode==6,] = "2"
full.t.aic = t.aic
full.t.bic = t.bic
full.t.sw = t.sw
t.aic = t.aic[-idx2,]
t.bic = t.bic[-idx2,]
t.sw = t.sw[-idx2,]
compare = tibble(full.short, full.t.sw, full.t.aic, full.t.bic)
```

```{r echo=F, warning=F, message=F}
# Data prepare
maxlag = 4
lambda.seq <- seq(0.5, 0, -0.0025)
#save.image(file = "sab_gdp.RData")
```


## Lasso 1  

$$ 
\begin{split} 
\Delta y_t &=y_{t-1} \\
&+ \Delta y_{t-1} + \Delta y_{t-2} + \Delta y_{t-3} + \Delta y_{t-4} \\
&+  I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} +  I(0)_{t-4} \\
&+  I(1)_{t-1} +  I(1)_{t-2} 
+  I(1)_{t-3} +  I(1)_{t-4} \\
&+  \Delta I(2)_{t-1} +  \Delta I(2)_{t-2} 
+  \Delta I(2)_{t-3} +  \Delta I(2)_{t-4}
\end{split} 
$$

```{r include=T, message=F, warning=F, echo=F}
# SW
lasso1.sw = lasso111(data = bpraw, tcode = t.sw)
trace1.sw = trace.plot.table(lasso1.sw$lasso, lambda.seq, lasso1.sw$lambda)
# AIC
lasso1.aic = lasso111(data = bpraw, tcode = t.aic)
trace1.aic = trace.plot.table(lasso1.aic$lasso, lambda.seq, lasso1.aic$lambda)
# BIC
lasso1.bic = lasso111(data = bpraw, tcode = t.bic)
trace1.bic = trace.plot.table(lasso1.bic$lasso, lambda.seq, lasso1.bic$lambda)

sum1 = sumtable(trace1.sw, trace1.aic, trace1.bic)
sum1 = cbind(I1 = c("", "", "AIC, BIC", "SW, AIC", "", "SW, AIC"), sum1)
colnames(sum1) = c("I(1)", "S & W", "AIC", "BIC")
pander(sum1, caption = "Lasso 1, non-zero coefficients")
```

## Lasso 2  

$$ 
\begin{split} 
\Delta y_t &=y_{t-1} \\
&+ \Delta y_{t-1} + \Delta y_{t-2} + \Delta y_{t-3} + \Delta y_{t-4} \\
&+  I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} +  I(0)_{t-4} \\
&+  \Delta I(1)_{t-1} +  \Delta I(1)_{t-2} 
+  \Delta I(1)_{t-3} +  \Delta I(1)_{t-4} \\
&+  \Delta^2 I(2)_{t-1} +  \Delta^2 I(2)_{t-2} 
+  \Delta^2 I(2)_{t-3} +  \Delta^2 I(2)_{t-4}
\end{split} 
$$


```{r include=T, message=F, warning=F, echo=F}
# SW
lasso2.sw = lasso222(data = bpraw, tcode = t.sw)
trace2.sw = trace.plot.table(lasso2.sw$lasso, lambda.seq, lasso2.sw$lambda)
# AIC
lasso2.aic = lasso222(data = bpraw, tcode = t.aic)
trace2.aic = trace.plot.table(lasso2.aic$lasso, lambda.seq, lasso2.aic$lambda)
# BIC
lasso2.bic = lasso222(data = bpraw, tcode = t.bic)
trace2.bic = trace.plot.table(lasso2.bic$lasso, lambda.seq, lasso2.bic$lambda)

sum2 = sumtable(trace2.sw, trace2.aic, trace2.bic)
colnames(sum2) = c("I(1)", "S & W", "AIC", "BIC")
pander(sum2, caption = "Lasso 2, non-zero coefficients")
```

## Lasso 3

$$ 
\begin{split} 
\Delta y_t &=y_{t-1} \\
&+ \Delta y_{t-1} + \Delta y_{t-2} + \Delta y_{t-3} + \Delta y_{t-4} \\
&+  I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} +  I(0)_{t-4} \\
&+  \Delta I(1)_{t-1} +  \Delta I(1)_{t-2} 
+  \Delta I(1)_{t-3} +  \Delta I(1)_{t-4} \\
&+  \Delta^2 I(2)_{t-1} +  \Delta^2 I(2)_{t-2} 
+  \Delta^2 I(2)_{t-3} +  \Delta^2 I(2)_{t-4} \\
&+ I(1)_{t-1} + \Delta I(2)_{t-1}
\end{split} 
$$

```{r include=T, message=F, warning=F, echo=F}
# SW
lasso3.sw = lasso333(data = bpraw, tcode = t.sw)
trace3.sw = trace.plot.table(lasso3.sw$lasso, lambda.seq, lasso3.sw$lambda)
# AIC
lasso3.aic = lasso333(data = bpraw, tcode = t.aic)
trace3.aic = trace.plot.table(lasso3.aic$lasso, lambda.seq, lasso3.aic$lambda)
# BIC
lasso3.bic = lasso333(data = bpraw, tcode = t.bic)
trace3.bic = trace.plot.table(lasso3.bic$lasso, lambda.seq, lasso3.bic$lambda)

sum3 = sumtable(trace3.sw, trace3.aic, trace3.bic)
sum3 = cbind(I1 = c("",  "AIC, BIC", "", "", "", "", ""), sum3)
colnames(sum3) = c("I(1)", "S & W", "AIC", "BIC")
pander(sum3, caption = "Lasso 3, non-zero coefficients")
```







