---
title: "SW and ADF test (Inflation)"
author: "Shuofan Zhang"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r, include=T, echo=F, warning=F, 	message = FALSE}
library("glmnet")
library("reshape2")
library("cellranger")
library("tidyverse")
library("knitr")
library("ggrepel")
library("pander")
library("kableExtra")
library("tseries")
library("urca")
library("zoo")
source("/Users/stanza/documents/github/lasso/fun.R")
load("/Users/stanza/documents/github/lasso/sab_gdp.RData")
```


* Stock and Watson twice differenced the CPI to make it stationary.

* ADF test with AIC/BIC conclude Inflation as I(0) with full set of observations.

* ADF test with AIC/BIC conclude Inflation as I(1) with the first 80% of observations.

```{r, echo=F, message=F, warning=F}
t.inf = t.sw
t.inf[short=="CPI"] = "1"
Inflation = diff(bpraw$CPI) %>% as.data.frame()
l = ceiling(dim(Inflation)[1]*0.8)
xx = as.yearqtr(1959 + seq(1, 223)/4) 
ggplot(data=Inflation, aes(x=xx, y=.)) + geom_line() + geom_vline(xintercept=xx[l], color="red")+
  xlab("time") + ylab("Inflation rate") + scale_x_continuous(breaks = seq(1959, 2014, by=5))
```

# Inflation as I(0)


```{r include=T, message=F, warning=F, echo=F}
#lasso1.sw = lasso111inf(data = bpraw, tcode = t.inf)
#trace1.sw = trace.plot.table(lasso1.sw$lasso, lambda.seq, lasso1.sw$lambda)
#lasso1.aic = lasso111inf(data = bpraw, tcode = t.aic)
#trace1.aic = trace.plot.table(lasso1.aic$lasso, lambda.seq, lasso1.aic$lambda)
#lasso1.bic = lasso111inf(data = bpraw, tcode = t.bic)
#trace1.bic = trace.plot.table(lasso1.bic$lasso, lambda.seq, lasso1.bic$lambda)
train1.sw = inf1train(data = bpraw, tcode = t.sw)
train1.aic = inf1train(data = bpraw, tcode = t.aic)
train1.bic = inf1train(data = bpraw, tcode = t.bic)
trace1.sw = trace.plot.table(train1.sw$lasso, lambda.seq, train1.sw$lambda)
trace1.aic = trace.plot.table(train1.aic$lasso, lambda.seq, train1.aic$lambda)
trace1.bic = trace.plot.table(train1.bic$lasso, lambda.seq, train1.bic$lambda)
sum1 = sumtable(trace1.sw, trace1.aic, trace1.bic)
mses1 = c("out-of sample MSE", round(train1.sw$mse, digits = 6), 
         round(train1.aic$mse, digits = 6), round(train1.bic$mse, digits = 6))
sum1 = rbind(mses1, sum1)
MSE1 = c(train1.sw$mse, train1.aic$mse, train1.bic$mse)
write.csv(sum1, "/Users/stanza/documents/github/lasso/inf1.csv")
```

```{r include=T, message=F, warning=F, echo=F}
#lasso2.sw = lasso222inf(data = bpraw, tcode = t.inf)
#trace2.sw = trace.plot.table(lasso2.sw$lasso, lambda.seq, lasso2.sw$lambda)
#lasso2.aic = lasso222inf(data = bpraw, tcode = t.aic)
#trace2.aic = trace.plot.table(lasso2.aic$lasso, lambda.seq, lasso2.aic$lambda)
#lasso2.bic = lasso222inf(data = bpraw, tcode = t.bic)
#trace2.bic = trace.plot.table(lasso2.bic$lasso, lambda.seq, lasso2.bic$lambda)
train2.sw = inf2train(data = bpraw, tcode = t.sw)
train2.aic = inf2train(data = bpraw, tcode = t.aic)
train2.bic = inf2train(data = bpraw, tcode = t.bic)
trace2.sw = trace.plot.table(train2.sw$lasso, lambda.seq, train2.sw$lambda)
trace2.aic = trace.plot.table(train2.aic$lasso, lambda.seq, train2.aic$lambda)
trace2.bic = trace.plot.table(train2.bic$lasso, lambda.seq, train2.bic$lambda)
sum2 = sumtable(trace2.sw, trace2.aic, trace2.bic)
mses2 = c("out-of sample MSE", round(train2.sw$mse, digits = 6), 
         round(train2.aic$mse, digits = 6), round(train2.bic$mse, digits = 6))
sum2 = rbind(mses2, sum2)
MSE2 = c(train2.sw$mse, train2.aic$mse, train2.bic$mse)
write.csv(sum2, "/Users/stanza/documents/github/lasso/inf2.csv")
```

```{r include=T, message=F, warning=F, echo=F}
#lasso3.sw = lasso333inf(data = bpraw, tcode = t.inf)
#trace3.sw = trace.plot.table(lasso3.sw$lasso, lambda.seq, lasso3.sw$lambda)
#lasso3.aic = lasso333inf(data = bpraw, tcode = t.aic)
#trace3.aic = trace.plot.table(lasso3.aic$lasso, lambda.seq, lasso3.aic$lambda)
#lasso3.bic = lasso333inf(data = bpraw, tcode = t.bic)
#trace3.bic = trace.plot.table(lasso3.bic$lasso, lambda.seq, lasso3.bic$lambda)
train3.sw = inf3train(data = bpraw, tcode = t.sw)
train3.aic = inf3train(data = bpraw, tcode = t.aic)
train3.bic = inf3train(data = bpraw, tcode = t.bic)
trace3.sw = trace.plot.table(train3.sw$lasso, lambda.seq, train3.sw$lambda)
trace3.aic = trace.plot.table(train3.aic$lasso, lambda.seq, train3.aic$lambda)
trace3.bic = trace.plot.table(train3.bic$lasso, lambda.seq, train3.bic$lambda)
sum3 = sumtable(trace3.sw, trace3.aic, trace3.bic)
mses3 = c("out-of sample MSE", round(train3.sw$mse, digits = 6), 
         round(train3.aic$mse, digits = 6), round(train3.bic$mse, digits = 6))
sum3 = rbind(mses3, sum3)
MSE3 = c(train3.sw$mse, train3.aic$mse, train3.bic$mse)
write.csv(sum3, "/Users/stanza/documents/github/lasso/inf3.csv")
mses = rbind(MSE1, MSE2, MSE3)
colnames(mses) = c("SW", "AIC", "BIC")
rownames(mses) = c("LASSO 1", "LASSO 2", "LASSO 3")
pander(mses, caption = "out-of-sample MSE, when the dependent variable is inflation")
```

* AIC chooses 3 lags for Inflation.

* The out-of-sample MSE of AR(3) is 2.266.

* The estimated AR(1) coefficient is 0.7565.

```{r include=T, message=F, warning=F, echo=F}
#ar(D.Inflation, aic = TRUE)
maxlag.inf = 3
Inflation = diff(bpraw$CPI)
y = lag0(maxlag = maxlag.inf, Inflation) %>% scale()
x = lags(maxlag = maxlag.inf, Inflation) 
l = ceiling(length(y)*0.8)
y.train = y[1:l]
y.test = y[(l+1):length(y)]
x.train = x[1:l, ] 
x.test = x[(l+1):dim(x)[1], ]
infols = lm(y.train ~ x.train)
cf = as.data.frame(lapply(infols$coefficients[-1], rep, dim(x.test)[1]))
fit.temp = cf * x.test
fit.inf = rowSums(fit.temp)
out.mse.inf = mean((y.test-fit.inf)^2)
#out.mse.inf
#ar(Inflation, order.max = 1)
rbind(adf.auto(Inflation, criteria = "AIC"), adf.auto(Inflation, criteria = "BIC"), adf.auto(y.train, criteria = "AIC"), adf.auto(y.train, criteria = "BIC")) %>% pander(caption="ADF test for Inflation")
```

# Inflation as I(1)

```{r include=T, message=F, warning=F, echo=F}
D.Inflation = diff(diff(bpraw$CPI)) %>% as.data.frame()
#ar(D.Inflation, order.max = 1)
l = ceiling(dim(D.Inflation)[1]*0.8)
xx = as.yearqtr(1959 + seq(2, 223)/4) 
ggplot(data=D.Inflation, aes(x=xx, y=.)) + geom_line() + geom_vline(xintercept=xx[l], color="red") +
  xlab("time") + ylab("Differenced Inflation") + scale_x_continuous(breaks = seq(1959, 2014, by=5))
```

```{r include=T, message=F, warning=F, echo=F}
t.aic.2 = t.aic
t.aic.2[short=="CPI"] = "2"
t.bic.2 = t.bic
t.bic.2[short=="CPI"] = "2"
#lasso1.sw2 = lasso111inf2(data = bpraw, tcode = t.sw)
#trace1.sw2 = trace.plot.table(lasso1.sw2$lasso, lambda.seq, lasso1.sw2$lambda)
#lasso1.aic2 = lasso111inf2(data = bpraw, tcode = t.aic.2)
#trace1.aic2 = trace.plot.table(lasso1.aic2$lasso, lambda.seq, lasso1.aic2$lambda)
#lasso1.bic2 = lasso111inf2(data = bpraw, tcode = t.bic.2)
#trace1.bic2 = trace.plot.table(lasso1.bic2$lasso, lambda.seq, lasso1.bic2$lambda)
train1.2.sw = inf1.2train(data = bpraw, tcode = t.sw)
train1.2.aic = inf1.2train(data = bpraw, tcode = t.aic.2)
train1.2.bic = inf1.2train(data = bpraw, tcode = t.bic.2)
trace1.2.sw = trace.plot.table(train1.2.sw$lasso, lambda.seq, train1.2.sw$lambda)
trace1.2.aic = trace.plot.table(train1.2.aic$lasso, lambda.seq, train1.2.aic$lambda)
trace1.2.bic = trace.plot.table(train1.2.bic$lasso, lambda.seq, train1.2.bic$lambda)
sum1.2 = sumtable(trace1.2.sw, trace1.2.aic, trace1.2.bic)
mses1.2 = c("out-of sample MSE", round(train1.2.sw$mse, digits = 6), 
         round(train1.2.aic$mse, digits = 6), round(train1.2.bic$mse, digits = 6))
sum1.2 = rbind(mses1.2, sum1.2)
MSE1.2 = c(train1.2.sw$mse, train1.2.aic$mse, train1.2.bic$mse)
write.csv(sum1.2, "/Users/stanza/documents/github/lasso/inf1_2.csv")
```

```{r include=T, message=F, warning=F, echo=F}
#lasso2.sw2 = lasso222inf(data = bpraw, tcode = t.sw)
#trace2.sw2 = trace.plot.table(lasso2.sw2$lasso, lambda.seq, lasso2.sw2$lambda)
#lasso2.aic2 = lasso222inf(data = bpraw, tcode = t.aic.2)
#trace2.aic2 = trace.plot.table(lasso2.aic2$lasso, lambda.seq, lasso2.aic2$lambda)
#lasso2.bic2 = lasso222inf(data = bpraw, tcode = t.bic.2)
#trace2.bic2 = trace.plot.table(lasso2.bic2$lasso, lambda.seq, lasso2.bic2$lambda)
train2.2.sw = inf2.2train(data = bpraw, tcode = t.sw)
train2.2.aic = inf2.2train(data = bpraw, tcode = t.aic.2)
train2.2.bic = inf2.2train(data = bpraw, tcode = t.bic.2)
trace2.2.sw = trace.plot.table(train2.2.sw$lasso, lambda.seq, train2.2.sw$lambda)
trace2.2.aic = trace.plot.table(train2.2.aic$lasso, lambda.seq, train2.2.aic$lambda)
trace2.2.bic = trace.plot.table(train2.2.bic$lasso, lambda.seq, train2.2.bic$lambda)
sum2.2 = sumtable(trace2.2.sw, trace2.2.aic, trace2.2.bic)
mses2.2 = c("out-of sample MSE", round(train2.2.sw$mse, digits = 6), 
         round(train2.2.aic$mse, digits = 6), round(train2.2.bic$mse, digits = 6))
sum2.2 = rbind(mses2.2, sum2.2)
MSE2.2 = c(train2.2.sw$mse, train2.2.aic$mse, train2.2.bic$mse)
write.csv(sum2.2, "/Users/stanza/documents/github/lasso/inf2_2.csv")
```

```{r include=T, message=F, warning=F, echo=F}
#lasso3.sw2 = lasso333inf(data = bpraw, tcode = t.sw)
#trace3.sw2 = trace.plot.table(lasso3.sw2$lasso, lambda.seq, lasso3.sw2$lambda)
#lasso3.aic2 = lasso333inf(data = bpraw, tcode = t.aic.2)
#trace3.aic2 = trace.plot.table(lasso3.aic2$lasso, lambda.seq, lasso3.aic2$lambda)
#lasso3.bic2 = lasso333inf(data = bpraw, tcode = t.bic.2)
#trace3.bic2 = trace.plot.table(lasso3.bic2$lasso, lambda.seq, lasso3.bic2$lambda)
train3.2.sw = inf3.2train(data = bpraw, tcode = t.sw)
train3.2.aic = inf3.2train(data = bpraw, tcode = t.aic.2)
train3.2.bic = inf3.2train(data = bpraw, tcode = t.bic.2)
trace3.2.sw = trace.plot.table(train3.2.sw$lasso, lambda.seq, train3.2.sw$lambda)
trace3.2.aic = trace.plot.table(train3.2.aic$lasso, lambda.seq, train3.2.aic$lambda)
trace3.2.bic = trace.plot.table(train3.2.bic$lasso, lambda.seq, train3.2.bic$lambda)
sum3.2 = sumtable(trace3.2.sw, trace3.2.aic, trace3.2.bic)
mses3.2 = c("out-of sample MSE", round(train3.2.sw$mse, digits = 6), 
         round(train3.2.aic$mse, digits = 6), round(train3.2.bic$mse, digits = 6))
sum3.2 = rbind(mses3.2, sum3.2)
MSE3.2 = c(train3.2.sw$mse, train3.2.aic$mse, train3.2.bic$mse)
write.csv(sum3.2, "/Users/stanza/documents/github/lasso/inf3_2.csv")
mses.2 = rbind(MSE1.2, MSE2.2, MSE3.2)
colnames(mses.2) = c("SW", "AIC", "BIC")
rownames(mses.2) = c("LASSO 1", "LASSO 2", "LASSO 3")
pander(mses.2, caption = "out-of-sample MSE, when the dependent variable is first-differenced inflation")
```


* AIC chooses 4 lags for the first-differenced inflation.

* The out-of-sample MSE of AR(4) is 1.969.

* The estimated AR(1) coefficient is -0.2728.


```{r include=T, message=F, warning=F, echo=F}
# differenced inflation
maxlag.d.inf = 4
D.Inflation = diff(Inflation)
y = lag0(maxlag = maxlag.d.inf, D.Inflation) %>% scale()
x = lags(maxlag = maxlag.d.inf, D.Inflation) 
l = ceiling(length(y)*0.8)
y.train = y[1:l]
y.test = y[(l+1):length(y)]
x.train = x[1:l, ] 
x.test = x[(l+1):dim(x)[1], ]
infols = lm(y.train ~ x.train)
cf = as.data.frame(lapply(infols$coefficients[-1], rep, dim(x.test)[1]))
fit.temp = cf * x.test
fit.inf = rowSums(fit.temp)
out.mse.Dinf = mean((y.test-fit.inf)^2)
#out.mse.Dinf
```




