---
title: "SW (GDP)"
author: "Shuofan Zhang"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r, include=T, echo=F, warning=F, 	message = FALSE}
load("/Users/stanza/documents/github/lasso/sab_gdp.RData")
source("/Users/stanza/documents/github/lasso/fun.R")
library("glmnet")
library("reshape2")
library("cellranger")
library("tidyverse")
library("knitr")
library("ggrepel")
library("pander")
library("kableExtra")
library("tseries")
library("urca")
library("zoo")
library("ggfortify")
```

## Data

Data: 145+1-5=141 series in total.

The "spread" series (difference between two I(1) series) are removed.

One of the I(1), "CP3FM" (short name: "Com Paper") was omitted in the original data set, now is added.

log() is done.

$y$ is GDP.

The earliest 80% of data were used to estimate the coefficients, the rest 20% were used to calculate the out-of-sample MSE.

```{r, include=T, echo=F, warning=F, 	message = FALSE}
D.GDP = diff(bpraw$GDP) %>% as.data.frame()
l = ceiling(dim(D.GDP)[1]*0.8)
xx = as.yearqtr(1959 + seq(1, 223)/4) 
ggplot(data=D.GDP, aes(x=xx, y=.)) + geom_line() + geom_vline(xintercept=xx[l], color="red") +
  xlab("Year") + ylab("GDP growth") + scale_x_continuous(breaks = seq(1959, 2014, by=5))
```

## ADF test  

Step 1, ADF test to the 146 original series.

Step 2, mark "I(0)" variables as "I(0)".

Step 3, ADF test to the first-differenced 146 series.

Step 4, check for contradictions, found "PCED_RecServices" in AIC.

Step 5, mark "I(1)" variable as "I(2)" (including "PCED_RecServices").

Step 6, mark the rest as "I(1)".

Step 7, repeat the above 6 steps for both "AIC" and "BIC".

## AR(p) modeL  

* The estimation method is OLS; lags are chosen by AIC.

## Lasso 1  

$$ 
\begin{split} 
\Delta y_t &=y_{t-1} \\
&+ \Delta y_{t-1} + \Delta y_{t-2} + \Delta y_{t-3} + \Delta y_{t-4} \\
&+  I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} +  I(0)_{t-4} \\
&+  I(1)_{t-1} +  I(1)_{t-2} 
+  I(1)_{t-3} +  I(1)_{t-4} \\
&+  \Delta I(2)_{t-1} +  \Delta I(2)_{t-2} 
+  \Delta I(2)_{t-3} +  \Delta I(2)_{t-4}
\end{split} 
$$

```{r include=T, message=F, warning=F, echo=F}
#lasso1.sw = lasso111gdp(data = bpraw, tcode = t.sw)
#trace1.sw = trace.plot.table(lasso1.sw$lasso, lambda.seq, lasso1.sw$lambda)
#lasso1.aic = lasso111gdp(data = bpraw, tcode = t.aic)
#trace1.aic = trace.plot.table(lasso1.aic$lasso, lambda.seq, lasso1.aic$lambda)
#lasso1.bic = lasso111gdp(data = bpraw, tcode = t.bic)
#trace1.bic = trace.plot.table(lasso1.bic$lasso, lambda.seq, lasso1.bic$lambda)
train1.sw = GDP1train(data = bpraw, tcode = t.sw)
train1.aic = GDP1train(data = bpraw, tcode = t.aic)
train1.bic = GDP1train(data = bpraw, tcode = t.bic)
trace1.sw = trace.plot.table(train1.sw$lasso, lambda.seq, train1.sw$lambda)
trace1.aic = trace.plot.table(train1.aic$lasso, lambda.seq, train1.aic$lambda)
trace1.bic = trace.plot.table(train1.bic$lasso, lambda.seq, train1.bic$lambda)
sum1 = sumtable(trace1.sw, trace1.aic, trace1.bic)
mses1 = c("out-of sample MSE", round(train1.sw$mse, digits = 6), 
         round(train1.aic$mse, digits = 6), round(train1.bic$mse, digits = 6))
sum1 = rbind(mses1, sum1)
write.csv(sum1, file="/Users/stanza/documents/github/lasso/gdp1.csv")
MSE1 = c(train1.sw$mse, train1.aic$mse, train1.bic$mse)
```

## Lasso 2  

$$ 
\begin{split} 
\Delta y_t &=y_{t-1} \\
&+ \Delta y_{t-1} + \Delta y_{t-2} + \Delta y_{t-3} + \Delta y_{t-4} \\
&+  I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} +  I(0)_{t-4} \\
&+  \Delta I(1)_{t-1} +  \Delta I(1)_{t-2} 
+  \Delta I(1)_{t-3} +  \Delta I(1)_{t-4} \\
&+  \Delta^2 I(2)_{t-1} +  \Delta^2 I(2)_{t-2} 
+  \Delta^2 I(2)_{t-3} +  \Delta^2 I(2)_{t-4}
\end{split} 
$$


```{r include=T, message=F, warning=F, echo=F}
#lasso2.sw = lasso222gdp(data = bpraw, tcode = t.sw)
#trace2.sw = trace.plot.table(lasso2.sw$lasso, lambda.seq, lasso2.sw$lambda)
#lasso2.aic = lasso222gdp(data = bpraw, tcode = t.aic)
#trace2.aic = trace.plot.table(lasso2.aic$lasso, lambda.seq, lasso2.aic$lambda)
#lasso2.bic = lasso222gdp(data = bpraw, tcode = t.bic)
#trace2.bic = trace.plot.table(lasso2.bic$lasso, lambda.seq, lasso2.bic$lambda)
train2.sw = GDP2train(data = bpraw, tcode = t.sw)
train2.aic = GDP2train(data = bpraw, tcode = t.aic)
train2.bic = GDP2train(data = bpraw, tcode = t.bic)
trace2.sw = trace.plot.table(train2.sw$lasso, lambda.seq, train2.sw$lambda)
trace2.aic = trace.plot.table(train2.aic$lasso, lambda.seq, train2.aic$lambda)
trace2.bic = trace.plot.table(train2.bic$lasso, lambda.seq, train2.bic$lambda)
sum2 = sumtable(trace2.sw, trace2.aic, trace2.bic)
mses2 = c("out-of sample MSE", round(train2.sw$mse, digits = 6), 
         round(train2.aic$mse, digits = 6), round(train2.bic$mse, digits = 6))
sum2 = rbind(mses2, sum2)
MSE2 = c(train2.sw$mse, train2.aic$mse, train2.bic$mse)
write.csv(sum2, file="/Users/stanza/documents/github/lasso/gdp2.csv")
```

## Lasso 3

$$ 
\begin{split} 
\Delta y_t &=y_{t-1} \\
&+ \Delta y_{t-1} + \Delta y_{t-2} + \Delta y_{t-3} + \Delta y_{t-4} \\
&+  I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} +  I(0)_{t-4} \\
&+  \Delta I(1)_{t-1} +  \Delta I(1)_{t-2} 
+  \Delta I(1)_{t-3} +  \Delta I(1)_{t-4} \\
&+  \Delta^2 I(2)_{t-1} +  \Delta^2 I(2)_{t-2} 
+  \Delta^2 I(2)_{t-3} +  \Delta^2 I(2)_{t-4} \\
&+ I(1)_{t-1} + \Delta I(2)_{t-1}
\end{split} 
$$

```{r include=T, message=F, warning=F, echo=F}
#lasso3.sw = lasso333gdp(data = bpraw, tcode = t.sw)
#trace3.sw = trace.plot.table(lasso3.sw$lasso, lambda.seq, lasso3.sw$lambda)
#lasso3.aic = lasso333gdp(data = bpraw, tcode = t.aic)
#trace3.aic = trace.plot.table(lasso3.aic$lasso, lambda.seq, lasso3.aic$lambda)
#lasso3.bic = lasso333gdp(data = bpraw, tcode = t.bic)
#trace3.bic = trace.plot.table(lasso3.bic$lasso, lambda.seq, lasso3.bic$lambda)
train3.sw = GDP3train(data = bpraw, tcode = t.sw)
train3.aic = GDP3train(data = bpraw, tcode = t.aic)
train3.bic = GDP3train(data = bpraw, tcode = t.bic)
trace3.sw = trace.plot.table(train3.sw$lasso, lambda.seq, train3.sw$lambda)
trace3.aic = trace.plot.table(train3.aic$lasso, lambda.seq, train3.aic$lambda)
trace3.bic = trace.plot.table(train3.bic$lasso, lambda.seq, train3.bic$lambda)
sum3 = sumtable(trace3.sw, trace3.aic, trace3.bic)
mses3 = c("out-of sample MSE", round(train3.sw$mse, digits = 6), 
         round(train3.aic$mse, digits = 6), round(train3.bic$mse, digits = 6))
sum3 = rbind(mses3, sum3)
write.csv(sum3, file="/Users/stanza/documents/github/lasso/gdp3.csv")
MSE3 = c(train3.sw$mse, train3.aic$mse, train3.bic$mse)
mses = rbind(MSE1, MSE2, MSE3)
colnames(mses) = c("SW", "AIC", "BIC")
rownames(mses) = c("LASSO 1", "LASSO 2", "LASSO 3")
pander(mses, caption = "out-of-sample MSE, when the dependent variable is inflation")
```

* AIC chooses 2 lags for the GDP growth.

* The out-of-sample MSE of AR(2) is 0.89.

* The estimated AR(1) coefficient is 0.3078.

```{r include=T, message=F, warning=F, echo=F}
D.GDP = diff(bpraw$GDP)
#ar(D.GDP, order.max = 1)
#ar(D.GDP, aic = TRUE)
maxlag.GDP = 2
y = lag0(maxlag = maxlag.GDP, D.GDP) %>% scale()
x = lags(maxlag = maxlag.GDP, D.GDP) 
l = ceiling(length(y)*0.8)
y.train = y[1:l]
y.test = y[(l+1):length(y)]
x.train = x[1:l, ] 
x.test = x[(l+1):dim(x)[1], ]
GDPols = lm(y.train ~ x.train)
cf = as.data.frame(lapply(GDPols$coefficients[-1], rep, dim(x.test)[1]))
fit.temp = cf * x.test
fit.GDP = rowSums(fit.temp)
out.mse.GDP = mean((y.test-fit.GDP)^2)
trace.interest(train1.sw$lasso, lambda.seq) 
#out.mse.GDP
```

The trace plot shows the paths of the four lags of the two interest rates that are I(1) variables: "3-month treasury bill (TB-3Mth)" and "3-Month AA Financial Commercial Paper Rate post 1997 … linked to XLI CP90 before 1997 (Com Paper)". 

Previously, when we had the difference of these two variables "CP_Tbill Spread" in the data, it was selected by the LASSO and the absolute value of its coefficient was the biggest among other non-zero coefficients. Since we have removed all the spreads in the data, I try to figure out what is happening to the two "parents" I(1) variables.
 
From the graph, we see:

1, when $\lambda=0$, which is the OLS, the magnitude and signs of the estimated coefficients all make sense, for example, lag-1 of TB-3Mth is positive and has a big coefficient, while lag-1 of Com Paper is negative and also has a big absolute coefficient. I think this may indicate that OLS works OK in terms of estimating co-integrating vectors.

2, Once $\lambda$ becomes non-zero, (in our case, the smallest non-zero $\lambda=0$ is 0.0025) seven out of eight of these variables are dropped immediately, which is a bit disappointing. 

Note: Although this graph only shows the trace of eight variables (two series with four lags), the regression was conducted with all 146 variables.

