---
title: "Data-driven Lambda for the LASSO"
author: "Shuofan Zhang"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

In this report, only the significant lags chosen by AR were included in the X matrix.

Since the plots for yr2, yr3, yr4 and yr5 are similar, so here I just plotted yr2.

The M.S.E of the training dataset increases with $\lambda$, so it is not feasible to choose a $\lambda$ according to the performance in training dataset.

Two solutions:

1, Same as before, separate the data set into training set (80%) and test set (20%); then choose $\lambda$ according to the smallest M.S.E in the test set directly and use it for the comparison.

2, Separate the data set into three sets, training (60%), validation (20%) and test (20%); then choose $\lambda$ according to the smallest M.S.E in the validation set, use the M.S.E from the test set for comparison.

(Continued from 12th Dec):

1, When we have high-dimentional data, and set *singular.ok = TRUE*, then run OLS with *lm* function in R, the x matrix will be truncated to have an appropriate number of columns. Example: n=100, p=200, the $100^{th}$-$200^{th}$ columns will be removed before running the regression. The actual data used in the regression will be n=100, p=99.

2, When we have high-dimentional data, and set $\lambda$ as a series of values including zero, then run LASSO with glmnet function in R, it gives non-zero estimations to every variables (when $\lambda=0$). In fact, the function does not allow $\lambda=0$ so it uses some very small value instead. (verified by data experiments as shown in the scatter plot, when the smallest lambda is smaller than 1E-6, the scatter plot almost reduce to a straight line, code check is ongoing)

3, When the data is not high-dimentional, the estimations of LASSO (when $\lambda=0$) becomes closer and closer to the OLS estimations with decreasing "thresh" values.

some references:

[How Correlations Influence Lasso Prediction](https://arxiv.org/pdf/1204.1605.pdf)

This paper argues correlation is not problematic for LASSO prediction, but it has influence on suitable $\lambda$. The higher the correlation, the smaller the tuning parameter.


\newpage

* LASSO 1 

The explanatory variables in this model are level I(1) variables, level I(0) variables, first-differenced I(2) variables and lags of the dependent variables.

```{r, include=T, echo=F, warning=F, 	message = FALSE}
library("glmnet")
library("reshape2")
library("cellranger")
library("tidyverse")
library("knitr")
library("ggrepel")
library("pander")
library("kableExtra")
library("tseries")
library("urca")
library("lmtest")
library("zoo")
load("/Users/stanza/documents/github/lasso/ng/ngdata.RData")
source("/Users/stanza/documents/github/lasso/fun.R")
maxlag = 25
lambda.seq <- seq(0.5, 0, -0.0025)
data = ngraw
tcode = t.ng
```

<!--
############### only sig lags #############################################################
x = x[, which( substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) == "lag1" |
                 substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) =="lag2" |
                 substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) =="lag3" |
                 substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) =="lag4" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag12" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag13" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag14" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag24" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag25" )]
############### only sig lags #############################################################
-->

```{r, include=T, echo=F, warning=F, 	message = FALSE}
# yr = yr2, LASSO 1
ytem = yr2[-1] %>% scale() # delete the first obs because of first differencing I(2) variables
y1 <- lag0(maxlag , ytem) %>% as.numeric()# dependent variable done
y.lags.1 <- lags(maxlag, ytem)
xtem1 = data[-1,]
xtem1[ , tcode=="2"] <- apply( data[ , tcode=="2"], 2, diff )
  colnames(xtem1)[tcode=="2"] = 
    paste0("D.", colnames(xtem1)[tcode=="2"])
xtem1 = lags(maxlag , xtem1) %>% scale() 
x1 = cbind(xtem1, y.lags.1)  # explanatory variables (2-223=222), done
x1.all = x1
############### only sig lags #############################################################
x1 = x1[, which( substr(colnames(x1), nchar(colnames(x1))-3, nchar(colnames(x1))) == "lag1" |
                 substr(colnames(x1), nchar(colnames(x1))-3, nchar(colnames(x1))) == "lag2" |
                 substr(colnames(x1), nchar(colnames(x1))-3, nchar(colnames(x1))) == "lag5" |
                 substr(colnames(x1), nchar(colnames(x1))-4, nchar(colnames(x1))) == "lag12" |
                 substr(colnames(x1), nchar(colnames(x1))-4, nchar(colnames(x1))) == "lag13")]
############### only sig lags #############################################################
OLS = lm(y1 ~ x1, singular.ok = TRUE)
coef0_LASSO = glmnet(x1, y1, alpha = 1, thresh = 1E-7, lambda = lambda.seq, maxit = 10^9)
lambda.non <- seq((0.5+10^-5), 10^-5, -0.0025)
coef_LASSO = glmnet(x1, y1, alpha = 1, thresh = 1E-7, lambda = lambda.non, maxit = 10^9)
coef0 = coef(coef0_LASSO)[-1,201]
coefnon = coef(coef_LASSO)[-1,201]
cc = cbind(as.vector(coef0),  as.vector(coefnon)) %>% as.data.frame()
ggplot(data = cc, aes(x=V1, y=V2)) + geom_point() + labs(caption = "Scatter plot of two sets of coefficients of LASSO when lambda equals to zero and 1E-5") + xlab("Lambda equals Zero") + ylab("Lambda equals 1E-5")
##################### compare OLS with LASSO end #############################################
yr2lasso1.1 = ddriven.l(x=x1, y=y1)
yr2lasso1.1$plot + labs(caption = "Path of M.S.E in two datasets against lambda, LASSO 1") + xlab("Lambda") + ylab("M.S.E")
yr2lasso1.2 = ddriven.l1l2(x=x1, y=y1)
yr2lasso1.2$plot + labs(caption = "Path of M.S.E in three datasets against lambda, LASSO 1") + xlab("Lambda") + ylab("M.S.E")
t1 = trace.plot.table(yr2lasso1.2$lasso, lambda.seq = lambda.seq, lambda.fix = yr2lasso1.2$lambda.valid)
# for the purpose of MCS, includ all lags 
yr2lasso1.1.all = ddriven.l(x=x1.all, y=y1)
yr2lasso1.2.all = ddriven.l1l2(x=x1.all, y=y1)
```

* LASSO 2  

The explanatory variables in this model are first-differenced I(1) variables, level I(0) variables, twice-differenced I(2) variables and lags of the dependent variables.

```{r, include=T, echo=F, warning=F, 	message = FALSE}
# yr = yr2, LASSO 2
ytem = yr2 %>% scale()
i0 = data[-(1:2) , tcode=="0"]
di1 = apply(data[,tcode=="1"], 2, diff)
di1 = di1[-1,]
colnames(di1) = paste0("D.", colnames(di1))
di2 = apply(data[,tcode=="2"], 2, diff)
d2i2 = apply(di2, 2, diff)
colnames(d2i2) = paste0("D2.", colnames(d2i2))
xtem2 = cbind(i0, di1, d2i2) %>% scale()
ytem = ytem[-(1:2)]
y.lags.2 <- lags(maxlag, ytem)
y2 = lag0(maxlag, ytem) %>% as.numeric()
xtem2 = lags(maxlag , xtem2)  # x lost 2 obs
x2 = cbind(y.lags.2, xtem2)
x2.all = x2
############### only sig lags #############################################################
x2 = x2[, which( substr(colnames(x2), nchar(colnames(x2))-3, nchar(colnames(x2))) == "lag1" |
                 substr(colnames(x2), nchar(colnames(x2))-3, nchar(colnames(x2))) == "lag2" |
                 substr(colnames(x2), nchar(colnames(x2))-3, nchar(colnames(x2))) == "lag5" |
                 substr(colnames(x2), nchar(colnames(x2))-4, nchar(colnames(x2))) == "lag12" |
                 substr(colnames(x2), nchar(colnames(x2))-4, nchar(colnames(x2))) == "lag13")]
############### only sig lags #############################################################
yr2lasso2.1 = ddriven.l(x=x2, y=y2)
yr2lasso2.1$plot + labs(caption = "Path of M.S.E in two datasets against lambda, LASSO 2") + xlab("Lambda") + ylab("M.S.E")
yr2lasso2.2 = ddriven.l1l2(x=x2, y=y2)
yr2lasso2.2$plot + labs(caption = "Path of M.S.E in three datasets against lambda, LASSO 2") + xlab("Lambda") + ylab("M.S.E")
t2 = trace.plot.table(yr2lasso2.2$lasso, lambda.seq = lambda.seq, lambda.fix = yr2lasso2.2$lambda.valid)
yr2lasso2.1.all = ddriven.l(x=x2.all, y=y2)
yr2lasso2.2.all = ddriven.l1l2(x=x2.all, y=y2)
```

* LASSO 3  

The explanatory variables in this model are first-differenced I(1) variables, level I(0) variables, twice-differenced I(2) variables,level I(1) variables, first-differenced I(2) variables and lags of the dependent variables.

```{r, include=T, echo=F, warning=F, 	message = FALSE}
# yr = yr2, LASSO 3
  ytem = yr2[-(1:2)] %>% scale() %>% as.matrix()
  y3 = lag0(maxlag, ytem) 
  y.lags.3 = lags(maxlag, ytem)
  i0 = data[ , tcode=="0"] # *
  i1 = data[,tcode=="1"] # *
  i2 = data[,tcode=="2"]
  di1 = apply(i1, 2, diff) # *
  di2 = apply(i2, 2, diff) # *
  d2i2 = apply(di2, 2, diff) # *
  colnames(di1) = paste0("D.", colnames(di1))
  colnames(di2) = paste0("D.", colnames(di2))
  colnames(d2i2) = paste0("D2.", colnames(d2i2))
  xtem3.0 = cbind(i0[-(1:2),] , di1[-1,] , d2i2) %>% scale()
  xtem3.0 = lags(maxlag, xtem3.0)
  xtem3.1 = cbind( i1[-(1:2),] , di2[-1,]) %>% scale()
  xtem3.1 = lag1(maxlag, xtem3.1)
  x3 = cbind(xtem3.0, xtem3.1, y.lags.3)
  x3.all = x3
############### only sig lags #############################################################
x3 = x3[, which( substr(colnames(x3), nchar(colnames(x3))-3, nchar(colnames(x3))) == "lag1" |
                 substr(colnames(x3), nchar(colnames(x3))-3, nchar(colnames(x3))) =="lag2" |
                 substr(colnames(x3), nchar(colnames(x3))-3, nchar(colnames(x3))) =="lag5" |
                 substr(colnames(x3), nchar(colnames(x3))-4, nchar(colnames(x3))) =="lag12" |
                 substr(colnames(x3), nchar(colnames(x3))-4, nchar(colnames(x3))) =="lag13")]
############### only sig lags #############################################################
yr2lasso3.1 = ddriven.l(x=x3, y=y3)
yr2lasso3.1$plot + labs(caption = "Path of M.S.E in two datasets against lambda, LASSO 3") + xlab("Lambda") + ylab("M.S.E")
yr2lasso3.2 = ddriven.l1l2(x=x3, y=y3)
yr2lasso3.2$plot + labs(caption = "Path of M.S.E in three datasets against lambda, LASSO 3") + xlab("Lambda") + ylab("M.S.E")
t3 = trace.plot.table(yr2lasso3.2$lasso, lambda.seq = lambda.seq, lambda.fix = yr2lasso3.2$lambda.valid)
yr2lasso3.1.all = ddriven.l(x=x3.all, y=y3)
yr2lasso3.2.all = ddriven.l1l2(x=x3.all, y=y3)
```

```{r, include=T, echo=F, warning=F, 	message = FALSE}
tt = sumtable(t1, t2, t3)
colnames(tt) = c("variable", "LASSO 1", "LASSO 2", "LASSO 3")
pander(tt, caption = "Variables selected by the LASSO with data driven lambda (valid) for yr2")
```

Graphs of the two sets of potential co-integrating vectors.

```{r, include=T, echo=F, warning=F, 	message = FALSE}
select(as.data.frame(x1), ces015.lag1, D.gmdcn.lag5) %>% mutate(WS = -0.06302 * ces015.lag1 -3 * (10^-6) * D.gmdcn.lag5) %>% mutate(time = 1:454) %>% melt(id="time") %>% ggplot(aes(x=time, y=value, color=variable)) + geom_line()  + facet_grid(rows = vars(variable)) + scale_x_continuous(breaks = c(seq(1, 450, by=90), 454), labels = as.yearmon(1964 + seq(0, 479)/12)[-(1:26)][c(seq(1, 450, by=90), 454)]) + xlab("Year") + ylab("Excess Bond Return (yr2)")

select(as.data.frame(x3), exrus.lag1, exruk.lag1) %>% mutate(WS = 0.002564 * exrus.lag1 - 0.002343 * exruk.lag1) %>% mutate(time = 1:453) %>% melt(id="time") %>% ggplot(aes(x=time, y=value, color=variable)) + geom_line()  + facet_grid(rows = vars(variable)) + scale_x_continuous(breaks = c(seq(1, 450, by=90), 453), labels = as.yearmon(1964 + seq(0, 479)/12)[-(1:27)][c(seq(1, 450, by=90), 453)]) + xlab("Year") + ylab("Excess Bond Return (yr2)")
```


Model Confidence Set 


```{r, include=T, echo=F, warning=F, 	message = FALSE}
library("MCS")
#data(Loss)
#MCS <- MCSprocedure(Loss=Loss[,1:5],alpha=0.2,B=5000,statistic='Tmax',cl=NULL)

load("/Users/stanza/documents/github/lasso/ng/res.RData")
load("/Users/stanza/documents/github/lasso/ng/res_all.RData")

Loss = cbind(L1.test = yr2lasso1.1$res, L1.valid = yr2lasso1.2$res, 
      L2.test = yr2lasso2.1$res, L2.valid = yr2lasso2.2$res, 
      L3.test = yr2lasso3.1$res, L3.valid = yr2lasso3.2$res,
      L1.test.all = yr2lasso1.1.all$res, L1.valid.all = yr2lasso1.2.all$res, 
      L2.test.all = yr2lasso2.1.all$res, L2.valid.all = yr2lasso2.2.all$res, 
      L3.test.all = yr2lasso3.1.all$res, L3.valid.all = yr2lasso3.2.all$res,
      AR = res.ar.yr2, AR.all = res.ar.yr2.all) %>% as.data.frame()
colnames(Loss) = c("L1.test", "L1.valid", "L2.test", "L2.valid", "L3.test", "L3.valid", "L1.test.all", "L1.valid.all", "L2.test.all", "L2.valid.all", "L3.test.all", "L3.valid.all","AR", "AR.all")

MCSprocedure(Loss=Loss,alpha=0.05, B=5000, statistic='Tmax',cl=NULL)

```
























