---
title: "Data-driven Lambda for the LASSO"
author: "Shuofan Zhang"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

In this report, only the significant lags chosen by AR were included in the X matrix.


Since the plots for yr2, yr3, yr4 and yr5 are similar, so here I just plotted yr2.


The M.S.E of the training dataset increases with $\lambda$, so it is not feasible to choose a $\lambda$ according to the performance in training dataset.


Two solutions:

1, Same as before, separate the data set into training set (80%) and test set (20%); then choose $\lambda$ according to the smallest M.S.E in the test set directly and use it for the comparison.

2, Separate the data set into three sets, training (60%), validation (20%) and test (20%); then choose $\lambda$ according to the smallest M.S.E in the validation set, use the M.S.E from the test set for comparison.

(Continued from 12th Dec):

1, When we have high-dimentional data, and set *singular.ok = TRUE*, then run OLS with *lm* function in R, the x matrix will be truncated to have an appropriate number of columns. Example: n=100, p=200, the $100^{th}$-$200^{th}$ columns will be removed before running the regression. The actual data used in the regression will be n=100, p=99.

2, When we have high-dimentional data, and set $\lambda$ as a series of values including zero, then run LASSO with glmnet function in R, it gives non-zero estimations to every variables (when $\lambda=0$). In fact, the function does not allow $\lambda=0$ so it uses some very small value instead. (verified by data experiments, code check is ongoing)

3, When the data is not high-dimentional, the estimations of LASSO (when $\lambda=0$) becomes closer and closer to the OLS estimations with decreasing "thresh" values.

some references:

[How Correlations Influence Lasso Prediction](https://arxiv.org/pdf/1204.1605.pdf)

This paper argues correlation is not problematic for LASSO prediction, but it has influence on suitable $\lambda$. The higher the correlation, the smaller the tuning parameter.



\newpage

* LASSO 1  


```{r, include=T, echo=F, warning=F, 	message = FALSE}
library("glmnet")
library("reshape2")
library("cellranger")
library("tidyverse")
library("knitr")
library("ggrepel")
library("pander")
library("kableExtra")
library("tseries")
library("urca")
library("lmtest")
library("zoo")
load("/Users/stanza/documents/github/lasso/ng/ngdata.RData")
source("/Users/stanza/documents/github/lasso/fun.R")
maxlag = 25
lambda.seq <- seq(0.5, 0, -0.0025)
```

<!--
############### only sig lags #############################################################
x = x[, which( substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) == "lag1" |
                 substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) =="lag2" |
                 substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) =="lag3" |
                 substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) =="lag4" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag12" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag13" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag14" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag24" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag25" )]
############### only sig lags #############################################################
-->




```{r, include=T, echo=F, warning=F, 	message = FALSE}
# yr = yr2, LASSO 1
yr = yr2[-1] %>% scale() # delete the first obs because of first differencing I(2) variables
data = ngraw
tcode = t.ng
y <- lag0(maxlag , yr) %>% as.numeric()# dependent variable done
y.lags <- lags(maxlag, yr)
xtem = ngraw[-1,]
xtem[ , tcode=="2"] <- apply( data[ , tcode=="2"], 2, diff )
  colnames(xtem)[tcode=="2"] = 
    paste0("D.", colnames(xtem)[tcode=="2"])
xtem = lags(maxlag , xtem) %>% scale() 
x = cbind(xtem, y.lags)  # explanatory variables (2-223=222), done
############### only sig lags #############################################################
x = x[, which( substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) == "lag1" |
                 substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) =="lag2" |
                 substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) =="lag5" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag12" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag13")]
############### only sig lags #############################################################
OLS = lm(y ~ x, singular.ok = TRUE)
coef0_LASSO = glmnet(x, y, alpha = 1, thresh = 1E-15, lambda = lambda.seq, maxit = 10^9)
lambda.non <- seq((0.5+10^-5), 10^-5, -0.0025)
coef_LASSO = glmnet(x, y, alpha = 1, thresh = 1E-15, lambda = lambda.non, maxit = 10^9)
coef0 = coef(coef0_LASSO)[-1,201]
coefnon = coef(coef_LASSO)[-1,201]
cc = cbind(as.vector(coef0),  as.vector(coefnon)) %>% as.data.frame()
ggplot(data = cc, aes(x=V1, y=V2)) + geom_point()
##################### compare OLS with LASSO end #############################################
yr2lasso1 = ddriven.l(x=x, y=y)
yr2lasso1$plot + labs(caption = "Path of M.S.E in two datasets against lambda, LASSO 1") + xlab("Lambda") + ylab("M.S.E")
yr2lasso1 = ddriven.l1l2(x=x, y=y)
yr2lasso1$plot + labs(caption = "Path of M.S.E in three datasets against lambda, LASSO 1") + xlab("Lambda") + ylab("M.S.E")
t1 = trace.plot.table(yr2lasso1$lasso, lambda.seq = lambda.seq, lambda.fix = yr2lasso1$lambda.valid)
```

* LASSO 2  

```{r, include=T, echo=F, warning=F, 	message = FALSE}
# yr = yr2, LASSO 2
yr = yr2 %>% scale()
data = ngraw
tcode = t.ng
i0 = data[-(1:2) , tcode=="0"]
di1 = apply(data[,tcode=="1"], 2, diff)
di1 = di1[-1,]
colnames(di1) = paste0("D.", colnames(di1))
di2 = apply(data[,tcode=="2"], 2, diff)
d2i2 = apply(di2, 2, diff)
colnames(d2i2) = paste0("D2.", colnames(d2i2))
xtem = cbind(i0, di1, d2i2) %>% scale()
yr = yr[-(1:2)]
y.lags <- lags(maxlag, yr)
y = lag0(maxlag, yr) %>% as.numeric()
xtem = lags(maxlag , xtem)  # x lost 2 obs
x = cbind(y.lags, xtem)
############### only sig lags #############################################################
x = x[, which( substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) == "lag1" |
                 substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) =="lag2" |
                 substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) =="lag5" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag12" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag13")]
############### only sig lags #############################################################
yr2lasso2 = ddriven.l(x=x, y=y)
yr2lasso2$plot + labs(caption = "Path of M.S.E in two datasets against lambda, LASSO 2") + xlab("Lambda") + ylab("M.S.E")
yr2lasso2 = ddriven.l1l2(x=x, y=y)
yr2lasso2$plot + labs(caption = "Path of M.S.E in three datasets against lambda, LASSO 2") + xlab("Lambda") + ylab("M.S.E")
t2 = trace.plot.table(yr2lasso2$lasso, lambda.seq = lambda.seq, lambda.fix = yr2lasso2$lambda.valid)
```

* LASSO 3  

```{r, include=T, echo=F, warning=F, 	message = FALSE}
# yr = yr2, LASSO 3
yr = yr2[-(1:2)] %>% scale() %>% as.matrix()
  y = lag0(maxlag, yr) 
  y.lags = lags(maxlag, yr)
  i0 = data[ , tcode=="0"] # *
  i1 = data[,tcode=="1"] # *
  i2 = data[,tcode=="2"]
  di1 = apply(i1, 2, diff) # *
  di2 = apply(i2, 2, diff) # *
  d2i2 = apply(di2, 2, diff) # *
  colnames(di1) = paste0("D.", colnames(di1))
  colnames(di2) = paste0("D.", colnames(di2))
  colnames(d2i2) = paste0("D2.", colnames(d2i2))
  xtem = cbind(i0[-(1:2),] , di1[-1,] , d2i2) %>% scale()
  xtem = lags(maxlag, xtem)
  xtem2 = cbind( i1[-(1:2),] , di2[-1,]) %>% scale()
  xtem2 = lag1(maxlag, xtem2)
  x = cbind(xtem, xtem2, y.lags)
############### only sig lags #############################################################
x = x[, which( substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) == "lag1" |
                 substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) =="lag2" |
                 substr(colnames(x), nchar(colnames(x))-3, nchar(colnames(x))) =="lag5" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag12" |
                 substr(colnames(x), nchar(colnames(x))-4, nchar(colnames(x))) =="lag13")]
############### only sig lags #############################################################
yr2lasso3 = ddriven.l(x=x, y=y)
yr2lasso3$plot + labs(caption = "Path of M.S.E in two datasets against lambda, LASSO 3") + xlab("Lambda") + ylab("M.S.E")
yr2lasso3 = ddriven.l1l2(x=x, y=y)
yr2lasso3$plot + labs(caption = "Path of M.S.E in three datasets against lambda, LASSO 3") + xlab("Lambda") + ylab("M.S.E")
t3 = trace.plot.table(yr2lasso3$lasso, lambda.seq = lambda.seq, lambda.fix = yr2lasso3$lambda.valid)
```

```{r, include=T, echo=F, warning=F, 	message = FALSE}
tt = sumtable(t1, t2, t3)
colnames(tt) = c("variable", "LASSO 1", "LASSO 2", "LASSO 3")
pander(tt, caption = "Variables selected by the LASSO with data driven lambda (valid) for yr2")
```







