---
title: "NG and ADF test (excess return)"
author: "Shuofan Zhang"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r, include=T, echo=F, warning=F, 	message = FALSE}
library("glmnet")
library("reshape2")
library("cellranger")
library("tidyverse")
library("knitr")
library("ggrepel")
library("pander")
library("kableExtra")
library("tseries")
library("urca")
source("/Users/stanza/documents/github/lasso/fun.R")
```

## Data

Data: 131-8=123 series in total.

The "spread" series (difference between two I(1) series) are removed.

log() is done.


```{r echo=F, message=F, warning=F}
maxlag = 4
ngraw <- read.csv("/Users/stanza/documents/github/lasso/ng/ngraw.csv")  
tcode <- read.csv("/Users/stanza/documents/github/lasso/ng/tcode.csv") 
short <- read.delim2("/Users/stanza/documents/github/lasso/ng/series.txt", header = FALSE,
                      stringsAsFactors=FALSE) %>% as.data.frame()
colnames(ngraw) = str_trim(short$V1)
ydata <- readxl::read_xls("/Users/stanza/documents/github/lasso/ng/yvar.xls")
ydata <- select(ydata, yr2, yr3, yr4, yr5)
ydata = ydata %>% mutate(time=(1:480))
dt.yr2 = lm(data = ydata, yr2 ~ time)$residuals
dt.yr3 = lm(data = ydata, yr3 ~ time)$residuals
dt.yr4 = lm(data = ydata, yr4 ~ time)$residuals
dt.yr5 = lm(data = ydata, yr5 ~ time)$residuals
fd.yr2 = diff(ydata$yr2)
fd.yr3 = diff(ydata$yr3)
fd.yr4 = diff(ydata$yr4)
fd.yr5 = diff(ydata$yr5)
yr2 = ydata$yr2
ngraw = ngraw[1:dim(ydata)[1] , ]
ngraw[, tcode==4 | tcode==5 | tcode==6] = log(ngraw[, tcode==4 | tcode==5 | tcode==6])
# keep full set of data before remove "spread"
full.raw <- ngraw
full.tcode <- tcode
full.short <- short
# remove spread
idx2 = which(colnames(ngraw) == "scp90" | 
               colnames(ngraw) == "sfygm3" | 
               colnames(ngraw) == "sfygm6" | 
               colnames(ngraw) == "sfygt1" | 
               colnames(ngraw) == "sfygt5" |
               colnames(ngraw) == "sfygt10" |
               colnames(ngraw) == "sfyaaac" |
               colnames(ngraw) == "sfybaac")
tcode <- tcode[-idx2, ]
short <- short[-idx2, ]
ngraw <- ngraw[ , -idx2]
```

## ADF test  

Step 1, ADF test to the 146 original series.

Step 2, mark "I(0)" variables as "I(0)".

Step 3, ADF test to the first-differenced 146 series.

Step 4, check for contradictions, found "PCED_RecServices" in AIC.

Step 5, mark "I(1)" variable as "I(2)" (including "PCED_RecServices").

Step 6, mark the rest as "I(1)".

Step 7, repeat the above 6 steps for both "AIC" and "BIC".


```{r, echo=F, message=F, warning=F}
# Investigation of the all variables
t.aic = data.frame(integration = character(length = 131), stringsAsFactors = FALSE)
t.bic = data.frame(integration = character(length = 131), stringsAsFactors = FALSE)
t.ng = data.frame(integration = character(length = 131), stringsAsFactors = FALSE)
adf.aic = adf.auto(full.raw, criteria = "AIC")
adf.bic = adf.auto(full.raw, criteria = "BIC")
t.aic[adf.aic$Conclusion=="I(0)",] = "0"
t.bic[adf.bic$Conclusion=="I(0)",] = "0"
diff.full.raw = apply(full.raw , 2, diff)
adf.aic.diff = adf.auto(diff.full.raw, criteria = "AIC")
adf.aic = cbind(adf.aic, adf.aic.diff)
adf.bic.diff = adf.auto(diff.full.raw, criteria = "BIC")
adf.bic = cbind(adf.bic, adf.bic.diff)
write.csv(adf.aic, "/Users/stanza/documents/github/lasso/ng/adf_aic_ng.csv")
write.csv(adf.bic, "/Users/stanza/documents/github/lasso/ng/adf_bic_ng.csv")
t.aic[adf.aic[,6]=="I(1)",] = "2"
t.bic[adf.bic[,6]=="I(1)",] = "2"
t.aic[t.aic=="",] = "1"
t.bic[t.bic=="",] = "1"
t.ng[full.tcode==1,] = "0"
t.ng[full.tcode==2,] = "1"
t.ng[full.tcode==4,] = "0"
t.ng[full.tcode==5,] = "1"
t.ng[full.tcode==6,] = "2"
full.t.aic = t.aic
full.t.bic = t.bic
full.t.ng = t.ng
t.aic = t.aic[-idx2,]
t.bic = t.bic[-idx2,]
t.ng = t.ng[-idx2,]
```

```{r, echo=F, message=F, warning=F, eval=T}
# Investigation of the all variables
inves.raw2 = full.raw
inves.tcode2 = full.tcode
ng2 = inves.raw2[,inves.tcode2==6]
#adf.auto(ng2, criteria = "AIC")
temp = apply(ng2, 2, diff)
temp = rbind(NA, temp)
inves.raw2[,inves.tcode2==6] = temp
adf.aic = adf.auto(inves.raw2, criteria = "AIC")
adf.bic = adf.auto(inves.raw2, criteria = "BIC")
ng = data.frame(Series = colnames(inves.raw2), ng = character(length = 131), 
                  stringsAsFactors = FALSE)
ng$ng[inves.tcode2==1] = "I(0)"
ng$ng[inves.tcode2==2] = "I(1)"
ng$ng[inves.tcode2==4] = "I(0)"
ng$ng[inves.tcode2==5] = "I(1)"
ng$ng[inves.tcode2==6] = "I(2)"
aabbss <- left_join(ng, adf.aic, by = "Series") %>% left_join(., adf.bic, by = "Series")
aabbss <- aabbss[,-4]
aabbss <- aabbss[,c(1,2,3,5,6,4,7)]
colnames(aabbss) = c("Series", "ng", "AIC", "BIC", "Type", "AIC lags", "BIC lags")
tt = aabbss[inves.tcode2==6,]
tt$AIC[tt$AIC=="I(1)"] <- "I(2)"
tt$BIC[tt$BIC=="I(1)"] <- "I(2)"
tt$AIC[tt$AIC=="I(0)"] <- "I(1)"
tt$BIC[tt$BIC=="I(0)"] <- "I(1)"
aabbss[inves.tcode2==6,] = tt
saveRDS(aabbss, file = "/Users/stanza/documents/github/lasso/ng/compare.rds")
```

```{r echo=F, warning=F, message=F}
# Data prepare
maxlag = 4
lambda.seq <- seq(0.5, 0, -0.0025)
lambda.seq.fd <- seq(0.2, 0, -0.001)
#save.image(file = "sab_gdp.RData")
```


## Lasso 1

I(2) is first differenced, others are original.

$$ 
\begin{split} 
y_t &= I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} +  I(0)_{t-4} \\
&+  I(1)_{t-1} +  I(1)_{t-2} 
+  I(1)_{t-3} +  I(1)_{t-4} \\
&+  \Delta I(2)_{t-1} +  \Delta I(2)_{t-2} 
+  \Delta I(2)_{t-3} +  \Delta I(2)_{t-4}
\end{split} 
$$

```{r include=T, message=F, warning=F, echo=F}
# ng
lasso1.ng = lasso111ng(data = ngraw, tcode = t.ng, yr = yr2)
trace1.ng = trace.plot.table(lasso1.ng$lasso, lambda.seq, lasso1.ng$lambda)
# AIC
lasso1.aic = lasso111ng(data = ngraw, tcode = t.aic, yr = yr2)
trace1.aic = trace.plot.table(lasso1.aic$lasso, lambda.seq, lasso1.aic$lambda)
# BIC
lasso1.bic = lasso111ng(data = ngraw, tcode = t.bic, yr = yr2)
trace1.bic = trace.plot.table(lasso1.bic$lasso, lambda.seq, lasso1.bic$lambda)

sum1 = sumtable(trace1.ng, trace1.aic, trace1.bic)
#sum1 = cbind(I1=c("NG", "NG, BIC", "AIC, BIC", "BIC",
#                  "ALL", "NG", "AIC, BIC", "ALL",
#                  "ALL", "NG, BIC", "ALL", "ALL"), sum1)
colnames(sum1) = c("", "NG", "AIC", "BIC")
write.csv(sum1, "/Users/stanza/documents/github/lasso/ng/lasso1.csv")
```

```{r include=T, message=F, warning=F, echo=F}
# detrend y data
lasso1.ng.dt = lasso111ng(data = ngraw, tcode = t.ng, yr = dt.yr2)
trace1.ng.dt = trace.plot.table(lasso1.ng.dt$lasso, lambda.seq, lasso1.ng.dt$lambda)
lasso1.aic.dt = lasso111ng(data = ngraw, tcode = t.aic, yr = dt.yr2)
trace1.aic.dt = trace.plot.table(lasso1.aic.dt$lasso, lambda.seq, lasso1.aic.dt$lambda)
lasso1.bic.dt = lasso111ng(data = ngraw, tcode = t.bic, yr = dt.yr2)
trace1.bic.dt = trace.plot.table(lasso1.bic.dt$lasso, lambda.seq, lasso1.bic.dt$lambda)

sum1.dt = sumtable(trace1.ng.dt, trace1.aic.dt, trace1.bic.dt)
#sum1.dt = cbind(I1=c("BIC", "BIC", "ALL", "ALL", "AIC, BIC", "ALL", "ALL", "AIC, BIC"), sum1.dt)
colnames(sum1.dt) = c("", "NG", "AIC", "BIC")
#pander(sum1.dt, caption = "Lasso 1, non-zero coefficients, detrended y series")
write.csv(sum1.dt, "/Users/stanza/documents/github/lasso/ng/lasso1_dt.csv")
```

```{r include=T, message=F, warning=F, echo=F}
# first-differenced y data
lasso1.ng.fd = lasso111ng(data = ngraw[-1,], tcode = t.ng, yr=fd.yr2)
trace1.ng.fd = trace.plot.table(lasso1.ng.fd$lasso, lambda.seq.fd, lasso1.ng.fd$lambda)
lasso1.aic.fd = lasso111ng(data = ngraw[-1,], tcode = t.aic, yr=fd.yr2)
trace1.aic.fd = trace.plot.table(lasso1.aic.fd$lasso, lambda.seq.fd, lasso1.aic.fd$lambda)
lasso1.bic.fd = lasso111ng(data = ngraw[-1,], tcode = t.bic, yr=fd.yr2)
trace1.bic.fd = trace.plot.table(lasso1.bic.fd$lasso, lambda.seq.fd, lasso1.bic.fd$lambda)

sum1.fd = sumtable(trace1.ng.fd, trace1.aic.fd, trace1.bic.fd)
colnames(sum1.fd) = c("", "NG", "AIC", "BIC")
#pander(sum1.fd, caption = "Lasso 1, non-zero coefficients, detrended y series")
write.csv(sum1.fd, "/Users/stanza/documents/github/lasso/ng/lasso1_fd.csv")
```

## Lasso 2

All stationary.

$$ 
\begin{split} 
y_t &= I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} +  I(0)_{t-4} \\
&+  \Delta I(1)_{t-1} +  \Delta I(1)_{t-2} 
+  \Delta I(1)_{t-3} +  \Delta I(1)_{t-4} \\
&+  \Delta^2 I(2)_{t-1} +  \Delta^2 I(2)_{t-2} 
+  \Delta^2 I(2)_{t-3} +  \Delta^2 I(2)_{t-4}
\end{split} 
$$


```{r include=T, message=F, warning=F, echo=F}
# ng
lasso2.ng = lasso222ng(data = ngraw, tcode = t.ng, yr=yr2)
trace2.ng = trace.plot.table(lasso2.ng$lasso, lambda.seq, lasso2.ng$lambda)
# AIC
lasso2.aic = lasso222ng(data = ngraw, tcode = t.aic, yr=yr2)
trace2.aic = trace.plot.table(lasso2.aic$lasso, lambda.seq, lasso2.aic$lambda)
# BIC
lasso2.bic = lasso222ng(data = ngraw, tcode = t.bic, yr=yr2)
trace2.bic = trace.plot.table(lasso2.bic$lasso, lambda.seq, lasso2.bic$lambda)
sum2 = sumtable(trace2.ng, trace2.aic, trace2.bic)
colnames(sum2) = c("", "NG", "AIC", "BIC")
#pander(sum2, caption = "Lasso 2, non-zero coefficients")
write.csv(sum2, "/Users/stanza/documents/github/lasso/ng/lasso2.csv")
```

```{r include=T, message=F, warning=F, echo=F}
# detrend y data
lasso2.ng.dt = lasso222ng(data = ngraw, tcode = t.ng, yr=dt.yr2)
trace2.ng.dt = trace.plot.table(lasso2.ng.dt$lasso, lambda.seq, lasso2.ng.dt$lambda)
lasso2.aic.dt = lasso222ng(data = ngraw, tcode = t.aic, yr=dt.yr2)
trace2.aic.dt = trace.plot.table(lasso2.aic.dt$lasso, lambda.seq, lasso2.aic.dt$lambda)
lasso2.bic.dt = lasso222ng(data = ngraw, tcode = t.bic, yr=dt.yr2)
trace2.bic.dt = trace.plot.table(lasso2.bic.dt$lasso, lambda.seq, lasso2.bic.dt$lambda)
sum2.dt = sumtable(trace2.ng.dt, trace2.aic.dt, trace2.bic.dt)
colnames(sum2.dt) = c("", "NG", "AIC", "BIC")
#pander(sum2.dt, caption = "Lasso 2, non-zero coefficients, detrended y series")
write.csv(sum2.dt, "/Users/stanza/documents/github/lasso/ng/lasso2_dt.csv")
```

```{r include=T, message=F, warning=F, echo=F}
# first-differenced y data
lasso2.ng.fd = lasso222ng(data = ngraw[-1,], tcode = t.ng, yr=fd.yr2)
trace2.ng.fd = trace.plot.table(lasso2.ng.fd$lasso, lambda.seq.fd, lasso2.ng.fd$lambda)
lasso2.aic.fd = lasso222ng(data = ngraw[-1,], tcode = t.aic, yr=fd.yr2)
trace2.aic.fd = trace.plot.table(lasso2.aic.fd$lasso, lambda.seq.fd, lasso2.aic.fd$lambda)
lasso2.bic.fd = lasso222ng(data = ngraw[-1,], tcode = t.bic, yr=fd.yr2)
trace2.bic.fd = trace.plot.table(lasso2.bic.fd$lasso, lambda.seq.fd, lasso2.bic.fd$lambda)
sum2.fd = sumtable(trace2.ng.fd, trace2.aic.fd, trace2.bic.fd)
colnames(sum2.fd) = c("", "NG", "AIC", "BIC")
#pander(sum2.fd, caption = "Lasso 2, non-zero coefficients, detrended y series")
write.csv(sum2.fd, "/Users/stanza/documents/github/lasso/ng/lasso2_fd.csv")
```

## Lasso 3

Combination of Lasso 1 and 2.

$$ 
\begin{split} 
y_t &=I(0)_{t-1} +  I(0)_{t-2} 
+  I(0)_{t-3} +  I(0)_{t-4} \\
&+  \Delta I(1)_{t-1} +  \Delta I(1)_{t-2} 
+  \Delta I(1)_{t-3} +  \Delta I(1)_{t-4} \\
&+  \Delta^2 I(2)_{t-1} +  \Delta^2 I(2)_{t-2} 
+  \Delta^2 I(2)_{t-3} +  \Delta^2 I(2)_{t-4} \\
&+ I(1)_{t-1} + \Delta I(2)_{t-1}
\end{split} 
$$

```{r include=T, message=F, warning=F, echo=F}
# ng
lasso3.ng = lasso333ng(data = ngraw, tcode = t.ng, yr=yr2)
trace3.ng = trace.plot.table(lasso3.ng$lasso, lambda.seq, lasso3.ng$lambda)
# AIC
lasso3.aic = lasso333ng(data = ngraw, tcode = t.aic, yr=yr2)
trace3.aic = trace.plot.table(lasso3.aic$lasso, lambda.seq, lasso3.aic$lambda)
# BIC
lasso3.bic = lasso333ng(data = ngraw, tcode = t.bic, yr=yr2)
trace3.bic = trace.plot.table(lasso3.bic$lasso, lambda.seq, lasso3.bic$lambda)
sum3 = sumtable(trace3.ng, trace3.aic, trace3.bic)
colnames(sum3) = c("", "NG", "AIC", "BIC")
#pander(sum3, caption = "Lasso 3, non-zero coefficients")
write.csv(sum3, "/Users/stanza/documents/github/lasso/ng/lasso3.csv")
```

```{r include=T, message=F, warning=F, echo=F}
# detrend y data
lasso3.ng.dt = lasso333ng(data = ngraw, tcode = t.ng, yr=dt.yr3)
trace3.ng.dt = trace.plot.table(lasso3.ng.dt$lasso, lambda.seq, lasso3.ng.dt$lambda)
lasso3.aic.dt = lasso333ng(data = ngraw, tcode = t.aic, yr=dt.yr3)
trace3.aic.dt = trace.plot.table(lasso3.aic.dt$lasso, lambda.seq, lasso3.aic.dt$lambda)
lasso3.bic.dt = lasso333ng(data = ngraw, tcode = t.bic, yr=dt.yr3)
trace3.bic.dt = trace.plot.table(lasso3.bic.dt$lasso, lambda.seq, lasso3.bic.dt$lambda)
sum3.dt = sumtable(trace3.ng.dt, trace3.aic.dt, trace3.bic.dt)
colnames(sum3.dt) = c("", "NG", "AIC", "BIC")
#pander(sum3.dt, caption = "Lasso 3, non-zero coefficients, detrended y series")
write.csv(sum3.dt, "/Users/stanza/documents/github/lasso/ng/lasso3_dt.csv")
```

```{r include=T, message=F, warning=F, echo=F}
# first-differenced y data
lasso3.ng.fd = lasso333ng(data = ngraw[-1,], tcode = t.ng, yr=fd.yr2)
trace3.ng.fd = trace.plot.table(lasso3.ng.fd$lasso, lambda.seq.fd, lasso3.ng.fd$lambda)
lasso3.aic.fd = lasso333ng(data = ngraw[-1,], tcode = t.aic, yr=fd.yr2)
trace3.aic.fd = trace.plot.table(lasso3.aic.fd$lasso, lambda.seq.fd, lasso3.aic.fd$lambda)
lasso3.bic.fd = lasso333ng(data = ngraw[-1,], tcode = t.bic, yr=fd.yr2)
trace3.bic.fd = trace.plot.table(lasso3.bic.fd$lasso, lambda.seq.fd, lasso3.bic.fd$lambda)
sum3.fd = sumtable(trace3.ng.fd, trace3.aic.fd, trace3.bic.fd)
colnames(sum3.fd) = c("", "NG", "AIC", "BIC")
#pander(sum3.fd, caption = "Lasso 3, non-zero coefficients, detrended y series")
write.csv(sum3.fd, "/Users/stanza/documents/github/lasso/ng/lasso3_fd.csv")
```


```{r include=F, message=F, warning=F, echo=F, eval=F}
description = data.frame(`Short description` = c("hsmw",
                                   "hsbsou",
                                   "lhel",
                                   "fygt1",
                                   "exrjan",
                                   "exrcan",
                                   "lhu680",
                                   "ces003",
                                   "fclbmc",
                                   "ces140",
                                   "ces011",
                                   "ces002",
                                   "hsbmw",
                                   "pmcp",
                                   "pwimsa") ,  
           `Long description` = c("Housing Starts:Midwest(Thous.U.)S.A." , 
  "Houses Authorized By Build. Permits:South(Thou.U.)S.A.",
  "Index Of Help-Wanted Advertising In Newspapers (1967=100;Sa)",
  "Interest Rate: U.S.Treasury Const Maturities,1-Yr.(percentage Per Ann,Nsa)",
  "Foreign Exchange Rate: Japan (Yen Per U.S.$)",
  "Foreign Exchange Rate: Canada (Canadian $ Per U.S.$)",
  "Unemploy.By Duration: Average(Mean)Duration In Weeks (Sa)",
  "Employees On Nonfarm Payrolls - Goods-Producing",
  "Wkly Rp Lg Com'l Banks:Net Change Com'l & Indus Loans(Bil$,Saar)",
  "Employees On Nonfarm Payrolls - Government",
  "Employees On Nonfarm Payrolls - Construction",
  "Employees On Nonfarm Payrolls: Total Private",
  "Houses Authorized By Build. Permits:Midwest(Thou.U.)S.A.",
  "Napm Commodity Prices Index (Percent)",
  "Producer Price Index:I ntermed Mat.Supplies & Components(82=100,Sa)"), stringsAsFactors = FALSE) 
#%>% pander(justify = "left" , split.table = Inf , split.cell = 80, style = "grid", caption = "Data description")
write.csv(description, "description.csv")
```

```{r include=F, message=F, warning=F, echo=F, eval=F}
t.bic[colnames(ngraw) == substr(sum3[1,1], 1, nchar(sum3[1,1])-5)]
t.bic[colnames(ngraw) == substr(sum3[2,1], 1, nchar(sum3[2,1])-5)]
t.bic[colnames(ngraw) == substr(sum3[3,1], 3, nchar(sum3[3,1])-5)]
t.bic[colnames(ngraw) == substr(sum3[4,1], 1, nchar(sum3[4,1])-5)]
t.bic[colnames(ngraw) == substr(sum3[5,1], 1, nchar(sum3[5,1])-5)]
t.bic[colnames(ngraw) == substr(sum3[6,1], 1, nchar(sum3[6,1])-5)]
t.bic[colnames(ngraw) == substr(sum3[7,1], 1, nchar(sum3[7,1])-5)]
t.bic[colnames(ngraw) == substr(sum3[8,1], 1, nchar(sum3[8,1])-5)]
t.bic[colnames(ngraw) == substr(sum3[9,1], 1, nchar(sum3[9,1])-5)]
t.bic[colnames(ngraw) == substr(sum3[10,1], 1, nchar(sum3[10,1])-5)]
t.bic[colnames(ngraw) == substr(sum3[11,1], 1, nchar(sum3[11,1])-5)]
```







